---
title: "Hybrids Survival GLMM"
author: "John Powers"
date: "Jan 18, 2018"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.table { width: 40%; }
</style>

Modified from hybridgerm.Rmd (Oct 2017)

#Guides
* Bolker et al. 2009. [Generalized linear mixed models: a practical guide for ecology and evolution](http://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(09)00019-6). Trends in Ecology & Evolution.
* Bolker. [GLMM FAQ](http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html)
* Bolker et al. 2011. GLMMs in action: gene-by-environment interaction in total fruit production of wild populations of *Arabidopsis thaliana*. [Part 1](http://www.cell.com/cms/attachment/601623/4742452/mmc1.pdf), [Part 2](http://www.cell.com/cms/attachment/601623/4742453/mmc2.pdf)
* Bolker. In: Ecological Statistics: Contemporary Theory and Application
     + [GLMM:worked examples](http://ms.mcmaster.ca/~bolker/R/misc/foxchapter/bolker_chap.html) 
     + [Reporting GLMM results](http://ms.mcmaster.ca/~bolker/misc/GLMM_results_report.pdf)
* Bolker et al. 2013. [Strategies for fitting nonlinear ecological models in R, AD Model Builder, and BUGS](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12044/full)
     + Bolker et al. 2012. [Owls example: a zero-inflated, generalized linear
mixed model for count data](https://groups.nceas.ucsb.edu/non-linear-modeling/projects/owls/WRITEUP/owls.pdf)
     + other [examples](https://groups.nceas.ucsb.edu/non-linear-modeling/) from Non-Linear Model Working Group
* [A Practical Guide to Mixed Models in R](http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html)
* Schielzeth and Nakagawa 2012. [Nested by design: model fitting and interpretation in a mixed model era](http://onlinelibrary.wiley.com/doi/10.1111/j.2041-210x.2012.00251.x/full). Methods in Ecology and Evolution.
* Stroup. [Rethinking the Analysis of Non-Normal Data in Plant and Soil Science](https://dl.sciencesocieties.org/publications/aj/articles/107/2/811)
* Elston et al. 2001. [Analysis of aggregation, a worked example: Numbers of ticks on red grouse chicks](https://www.researchgate.net/publication/11945272_Analysis_of_aggregation_a_worked_example_Numbers_of_ticks_on_red_grouse_chicks)
* Annika Nelson. [Ant GLMM Example](http://rpubs.com/annikanelson/GLMM_Example)
* R package *pscl* vignette: [Regression Models for Count Data in R](https://cran.r-project.org/web/packages/pscl/vignettes/countreg.pdf)

#Purpose
Identify reproductive barriers between two sympatric moth-fflinated plant species, *Schiedea kaalae* and *S. hookeri* by fitting a generalized linear mixed model (GLMM).

In the experimental design, the following **crosstypes** were made:

* within species, **between** population (may show outbreeding depression or heterosis)
* within species, **within** populations (may show inbreeding depression)
* **hybrids** between species (indicates species barrier from fflination to seed production)

In this analysis the response variable is the fflen viability of each cross. Other barriers (hybrid survival, flowering) could be analyzed in a similar framework, with appropriate changes to the underlying distribution. 

Fixed effects:

* **crosstype** - hybrids, within population, between populations
* **species** - species of the maternal plant that produced the Viability

Potential random effects:

* **mompop** - maternal plant population
* **mompid** - maternal plant, specified by its population and ID
* **dadpop** - paternal plant population

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning= F, comment="  ", cache=T)
library(lme4)
library(MASS)
library(car)
library(glmmADMB)
library(bbmle)
library(lsmeans)
library(vcd)
library(countreg)

source("glmm_funs.R")

#Visualization
library(coefplot2)
library(ggplot2); theme_set(theme_gray())
library(lattice)
library(plyr)
library(beanplot)
library(knitr)
library(RColorBrewer)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

#Data Import

```{r data}
ff <- read.table("firstflower6.csv", header=T, sep="\t")
ff$firstflower.date[ff$firstflower.date==""] <- NA
ff$firstflower.date[ff$use.firstflower!="yes"] <- NA
ff$firstflower.date <- as.Date(ff$firstflower.date)
ff <- ff[ff$crossid!=107,] ##FIND OUT WHAT CROSS THIS IS
ff$alive[ff$use.alive.flowered!="yes"] <- NA
ff <- ff[!is.na(ff$alive),]

ff$firstflower <- as.integer(round(difftime(ff$firstflower.date, "2016-03-01")))
ff$alive <- ff$alive=="yes"


crosses <- read.table("hybrids.csv", header=T, sep="\t", colClasses=c(mompop="factor", dadpop="factor"))

crosscol <- c("green","blue","orange","red")

#treat populations as factors
ff$mompop <- crosses$mompop[match(ff$crossid, crosses$crossid)]
ff$momid <- crosses$momid[match(ff$crossid, crosses$crossid)]
ff$species <- crosses$momsp[match(ff$crossid, crosses$crossid)]
ff$dadpop <- crosses$dadpop[match(ff$crossid, crosses$crossid)]
ff$dadid <- crosses$dadid[match(ff$crossid, crosses$crossid)]
ff$dadsp <- crosses$dadsp[match(ff$crossid, crosses$crossid)]
ff$crosstype <- crosses$crosstype[match(ff$crossid, crosses$crossid)]
ff$cross <- crosses$cross[match(ff$crossid, crosses$crossid)]

#rename crosstype codes
ff$crosstype <- factor(ff$crosstype, levels=c("between", "within", "hybrid"))
#made "between" the first reference level to facilitate comparison between outcrossing populations and hybridizing species 

ff$mompop <- sapply(ff$mompop, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"))
ff$dadpop <- sapply(ff$dadpop, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"))

#define interactions
ff <- within(ff, sxc <- interaction(species,crosstype))
ff <- within(ff, sxcxm <- interaction(species,crosstype,mompop,momid))
ff <- within(ff, mompid <- as.factor(paste(mompop,momid,sep=".")))
ff <- within(ff, dadpid <- as.factor(paste(dadpop,dadid,sep=".")))
ff <- within(ff, smompop <- as.factor(paste(species,mompop,sep="")))

#check final structure
str(ff)
```

#Data Inspection

##Replication
The sample sizes are unbalanced at all levels, including maternal population: 
```{r reps}
reptab <- with(ff, table(smompop,crosstype))
mosaic(reptab, pop=F)
labeling_cells(text = reptab, margin = 0)(reptab)
```

Replication is low for some **within**-population crosses. The replication is even lower for each maternal plant, so we need to be wary of estimates when subsetting at this level:

```{r reps_mpid}
with(ff, kable(table(mompid,crosstype)))
```

##Overall data distribution

To identify the best-fitting distribution, we make quantile-quantile plots of the raw data against various distributions. The more points within the confidence interval envelopes, the better the fit. Later, we present quantile-quantile plots of the model residuals to assess model fit.

##Fixed effects
Effects and interactions in these plots are simply given by the mean, which may be unduly influenced by high values.
```{r maineff}
intplot <- ggplot(ff,aes(fill=factor(alive))) + geom_bar(position="fill")  + labs(y="Proportion") + guides(fill=guide_legend(title="Survived")) + facet_grid(~species)
intplot + aes(x=crosstype) 
```

##Random effects

### Maternal population
```{r randeff_mp}
intplot + aes(x=mompop) 
```

### Maternal plant
```{r randeff_mpid}
intplot + aes(x=mompid)
```

### Paternal population
```{r randeff_dp}
intplot + aes(x=dadpop)
```

#Models
We constructed the following models with the package *glmmADMB*. They all have the same fixed effects, **species x crosstype**, and response variable, **alive**

* X  = standard GL(M)M

| distribution, Random Effects: | None | Maternal plant | Maternal population |
|-------------------------------|------|----------------|---------------------|
| normal (norm) | X | X | X |

```{r models, echo=T, cache=T}
sc.bin          <- glm(alive~species*crosstype, data=ff, family="binomial")
sc.mix.mompid.bin         <- glmer(alive~species*crosstype + (1|mompid), data=ff, family="binomial")
sc.mix.mompop.bin    <- glmer(alive~species*crosstype + (1|mompop), data=ff, family="binomial")
sc.mix.momdadpid.bin         <- glmer(alive~species*crosstype + (1|mompid) + (1|dadpid), data=ff, family="binomial")
```

#Model comparison
##AIC
We will use the Aikake Information Criterion to pick the model the best fits the data, penalized by the number of parameters. Differences of 2 units are significant.


```{r aic}
#AICtab(sc.b, sc.mix.mompop.b, sc.mix.mompid.b)
sc.names <- c("sc.bin", "sc.mix.mompid.bin", "sc.mix.mompop.bin","sc.mix.momdadpid.bin")
sc.list <- sapply(sc.names, get, USE.NAMES=T)
sc.AIC <- ICtab(sc.list,mnames=sc.names,type="AIC", base=T, delta=F) # for AICc, nobs=nobs(sc.list[[1]])
class(sc.AIC)<-"data.frame"
all.names <- c(sc.names)
all.list <- sapply(all.names, get, USE.NAMES=T)
all.AIC <- dfun(rbind(sc.AIC))
all.AIC <- all.AIC[order(all.AIC$dAIC),]
kable(all.AIC, format.arg=list(digits=3))
```

The best-fiting model is a model with the following components:

* 
     + response: **alive**
     + fixed effects: **species**, **crosstype**, **species x crosstype**
     + random effect: 


##Coefficients
The coefficients estimated for each model agree qualitatively.


```{r coef, fig.asp=1}
sc.log.names <- sc.names
sc.log <- sapply(sc.log.names, get, USE.NAMES=T)

coefplot2(sc.log, legend.x="topright",legend=T,legend.args=list(cex=0.8, xpd=T, inset=c(-0.1,0)), col.pts=sample(gg_color_hue(length(sc.log.names))), spacing=0.05, lwd.2=2, lwd.1=4, intercept=T)
```

#Inference
We chose the model with nearly the best (lowest) AIC, to carry out inference tests and parameter estimation.

##Description


```{r pick}
mod <- sc.mix.momdadpid.bin
print(mod)
```

##Test significance of interaction
By dropping it from the model and performing a likelihood-ratio test, we see that the **species x crosstype** interaction is not significant.


```{r testint}
sxc.chisq <- drop1(mod, test="Chisq") #load from file
dfun(sxc.chisq)
```

##Model summary
The model estimated the following parameters, with individual parameter significance determined by the Wald z-test, and fixed effect significance determined by analysis of deviance Wald test.


```{r summ}
summary(mod)
Anova(mod, type=3)
```

##Least square means
The least square means procedure can generate predictor estimates of each type, and give their significance groupings with a post-hoc Tukey test. *S. hookeri*-produced hybrids produce less Viability than either crosses between or within *S. hookeri* populations. The other differences are not significant, but remember that the fixed effect of hybrid (vs. between) was significant (model summary).


```{r lsmeans, fig.asp=1, fig.width=5}
#Count
rg <- ref.grid(mod)
#summary(rg)
sxc.lsm <- lsmeans(rg, ~ crosstype*species)
plot(sxc.lsm)
cld.mod <- cld(sxc.lsm, Letters=letters) #tukey letterings
library(boot)
cld.mod$response <- inv.logit(cld.mod$lsmean)
cld.mod$uSE <- inv.logit(cld.mod$lsmean+cld.mod$SE)
cld.mod$lSE <- inv.logit(cld.mod$lsmean-cld.mod$SE)
options(digits=4)
cld.mod[rev(order(cld.mod$species, cld.mod$crosstype)),]

H.wb <-  with(cld.mod[cld.mod$species=="hook",], response[crosstype=="within"]/response[crosstype=="between"] - 1)
K.wb <- with(cld.mod[cld.mod$species=="kaal",], response[crosstype=="within"]/response[crosstype=="between"] - 1)
K.H <- with(cld.mod[cld.mod$crosstype=="between",], response[species=="kaal"]/response[species=="hook"] - 1)
maxsp <- ifelse(K.H>0, "kaal","hook")
minsp <- ifelse(K.H<0, "kaal","hook")
maxresp <- with(cld.mod, response[species==maxsp & crosstype=="between"])
minresp <- with(cld.mod, response[species==minsp & crosstype=="between"])
HK.resp <-  with(cld.mod, response[species=="hook" & crosstype=="hybrid"])
KH.resp <-  with(cld.mod, response[species=="kaal" & crosstype=="hybrid"])
HK.int <-   with(cld.mod, ifelse(HK.resp > minresp & HK.resp < maxresp, (HK.resp-minresp)/(maxresp-minresp), 
                                 ifelse(HK.resp < minresp, HK.resp/minresp-1, HK.resp/maxresp-1)))
KH.int <-   with(cld.mod, ifelse(KH.resp > minresp & KH.resp < maxresp, (KH.resp-minresp)/(maxresp-minresp), 
                                 ifelse(KH.resp < minresp, KH.resp/minresp-1, KH.resp/maxresp-1)))

intermed <- (minresp + maxresp) / 2

round(c(H.wb,K.wb,K.H,HK.int,KH.int),2)

ggplot(as.data.frame(cld.mod), aes(y=response, x=relevel(crosstype, "within"), fill=species)) +
  geom_col(position=position_dodge2()) +
  geom_linerange(aes(ymin=lSE, ymax=uSE), position=position_dodge(0.9)) +
  labs(x="", y="Survival",fill="Maternal species") +
  scale_fill_manual(labels = c("S. hookeri  ", "S. kaalae  "), values=brewer.pal(name="Set1", n=3)[c(3,2)]) +
  scale_x_discrete(labels = c("Intrapopulation", "Interpopulation", "Hybrid")) +
  geom_text(aes(label=.group), position=position_dodge(0.9), hjust=0, vjust=-1) +
  scale_y_continuous(expand = expand_scale(mult=c(0,0.05)), breaks = scales::pretty_breaks(n = 5), limits=c(0,1)) +
  theme_classic() + theme(legend.text=element_text(face="italic", size=rel(1)), legend.position="bottom", axis.text = element_text(colour="black", size=rel(1)), text=element_text(size=14), axis.ticks.x = element_blank()) + geom_segment(aes(x=2.5, y=intermed, xend=3.5, yend=intermed))
```


