---
title: "Hybrids Vegetative Biomass GLMM"
author: "John Powers"
date: "May 14, 2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.table { width: 40%; }
</style>

Modified from hybridInflo Biomass.Rmd (April 2017)

# Purpose
Identify reproductive barriers between two sympatric moth-pollinated plant species, *Schiedea kaalae* and *S. hookeri* by fitting a generalized linear mixed model (GLMM).

In the experimental design, the following **crosstypes** were made:

* within species, **between** population (may show outbreeding depression or heterosis)
* within species, **within** populations (may show inbreeding depression)
* **hybrids** between species (indicates species barrier from pollination to seed production)

In this analysis the response variable is the biomass of the offspring produced by each cross. Other barriers (hybrid survival, flowering) could be analyzed in a similar framework, with appropriate changes to the underlying distribution. 

Fixed effects:

* **crosstype** - hybrids, within population, between populations
* **species** - species of the maternal plant that produced the Vegetative Biomass

Potential random effects:

* **mompop** - maternal plant population
* **mompid** - maternal plant, specified by its population and ID
* **dadpop** - paternal plant population

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning= F, comment="  ", cache=T)
library(lme4)
library(MASS)
library(car)
library(glmmADMB)
library(bbmle)
library(lsmeans)
library(vcd)
library(countreg)

source("glmm_funs.R")

#Visualization
library(coefplot2)
library(ggplot2); theme_set(theme_gray())
library(lattice)
library(plyr); library(dplyr)
library(beanplot)
library(knitr)
library(RColorBrewer)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

# Data Import

```{r data}
ib <- read.table("vegbiomass.csv", header=T, sep="\t", 
                 colClasses=c(collect.date="Date", weigh.date="Date"))

crosses <- read.table("hybrids.csv", header=T, sep="\t", colClasses=c(mompop="factor", dadpop="factor"))

crosscol <- c("green","blue","orange","red")

#treat populations as factors
ib$mompop <- crosses$mompop[match(ib$crossid, crosses$crossid)]
ib$momid <- crosses$momid[match(ib$crossid, crosses$crossid)]
ib$species <- crosses$momsp[match(ib$crossid, crosses$crossid)]
ib$dadpop <- crosses$dadpop[match(ib$crossid, crosses$crossid)]
ib$dadid <- crosses$dadid[match(ib$crossid, crosses$crossid)]
ib$dadsp <- crosses$dadsp[match(ib$crossid, crosses$crossid)]
ib$crosstype <- crosses$crosstype[match(ib$crossid, crosses$crossid)]
ib$cross <- crosses$cross[match(ib$crossid, crosses$crossid)]
ib <- ib[!is.na(ib$crosstype),] #missing information for 2 samples

#rename crosstype codes
ib$crosstype <- factor(ib$crosstype, levels=c("between", "within", "hybrid"))
#made "between" the first reference level to facilitate comparison between outcrossing populations and hybridizing species 

ib$mompop <- sapply(ib$mompop, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"))
ib$dadpop <- sapply(ib$dadpop, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"))

#define interactions
ib <- within(ib, sxc <- interaction(species,crosstype))
ib <- within(ib, sxcxm <- interaction(species,crosstype,mompop,momid))
ib <- within(ib, mompid <- as.factor(paste(mompop,momid,sep=".")))
ib <- within(ib, dadpid <- as.factor(paste(dadpop,dadid,sep=".")))
ib <- within(ib, smompop <- as.factor(paste(species,mompop,sep="")))

ib$collect.date <- round(difftime(ib$collect.date, "2016-03-10"))

#check final structure
str(ib)


```

# Data Inspection

```{r massdist, warning=FALSE}
library(ggplot2)
ggplot(ib, aes(x=log10(mass), fill=cross, color=cross)) + geom_density(alpha=0.1, bw=0.06) + scale_fill_manual(values=crosscol) +scale_color_manual(values=crosscol)
```

## Replication
The sample sizes are unbalanced at all levels, including maternal population: 
```{r reps}
reptab <- with(ib, table(smompop,crosstype))
mosaic(reptab, pop=F)
labeling_cells(text = reptab, margin = 0)(reptab)
```

Replication is low for some **within**-population crosses. The replication is even lower for each maternal plant, so we need to be wary of estimates when subsetting at this level:

```{r reps_mpid}
with(ib, kable(table(mompid,crosstype)))
```

## Overall data distribution

To identify the best-fitting distribution, we make quantile-quantile plots of the raw data against various distributions. The more points within the confidence interval envelopes, the better the fit. Later, we present quantile-quantile plots of the model residuals to assess model fit.

```{r dists}
#QQ plots against various distributions
set.seed(1)
par(mfrow=c(1,3))
normal <- fitdistr(log10(ib$mass+1), "normal")
qqp(log10(ib$mass+1), "norm", main="Normal")
lognormal <- fitdistr(ib$mass+1, "lognormal")
qqp(ib$mass+1, "lnorm", main="Log Normal")
#pois <- fitdistr(ib$mass+1, "Poisson")
#qqp(ib$mass, "pois", pois$estimate, main="Poisson")
#nbinom <- fitdistr(ib$mass+1, "Negative Binomial") 
#qqp(ib$mass+1, "nbinom", size = nbinom$estimate[[1]], mu=nbinom$estimate[[2]], main="Negative Binomial")
gamma <- fitdistr(ib$mass+1, "gamma")
qqp(ib$mass+1, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]], main="Gamma")
```

## Distributions by fixed factors
```{r hst_fixed}
ggplot(ib, aes(x = mass, fill=species)) +
  geom_histogram(data=subset(ib,species == "hook"), aes(y=-..density..),binwidth=2)+
  geom_histogram(data=subset(ib,species == "kaal"), aes(y= ..density..),binwidth=2)+
  coord_flip() + facet_grid(~crosstype) + labs(y="Histogram", x="Vegetative biomass (g)") + 
  scale_fill_manual("Maternal species", labels = c("S. hookeri  ", "S. kaalae  "), values=brewer.pal(name="Set1", n=3)[c(3,2)]) + theme_classic() + theme(legend.text=element_text(face="italic", size=rel(1)), legend.position="bottom", axis.text = element_text(colour="black", size=rel(1)), text=element_text(size=14), axis.ticks.x = element_blank())
```

## Distributions by random factors
```{r hst_rand_mpid}
ggplot(aes(y=mass, x=mompid, color=crosstype), data=ib) + geom_count(alpha=0.8) + coord_flip() + labs(x="Maternal plant", y="Mass")
```

## Homogeneity of variances across subsets
Our mixed model uses one parameter to capture random effect variance, which is assumed to be homogeneous. Plotting on a log scale should uncouple variances from means to assess this visually. Subsets are **species** \* **crosstype** \* **maternal plant**.

Subset variances are not homogeneous:
```{r varhom}
ggplot(aes(y=log10(mass+1), x=sxcxm, color=crosstype), data=ib) + geom_boxplot() + coord_flip() + labs(y="ln(Vegetative Biomass + 1)",x="Subsets")
```


## Subset mean-variance relationship
Various distributions make different assumptions about the mean-variance (µ-Var) ratio. 

```{r meanvar}
grpVars <- with(ib, tapply(mass, list(sxcxm), var))
grpMeans <- with(ib, tapply(mass, list(sxcxm), mean))
grpCounts <- with(ib, tapply(mass, list(sxcxm), length))
#set weight=grpCounts to weight loess by sample sizes
ggplot(na.omit(data.frame(grpMeans,grpVars,grpCounts)),
       aes(x=grpMeans,y=grpVars, weight=1))+geom_point(aes(size=grpCounts))+
  guides(colour=guide_legend(title="Fit"),size=guide_legend(title="Sample size")) + labs(x="Subset Mean", y="Subset Variance") + labs(subtitle="Subset: species*crosstype*mompid")
```

## Fixed effects
Effects and interactions in these plots are simply given by the mean, which may be unduly influenced by high values.
```{r maineff}
intplot <- ggplot(ib,aes(x=crosstype,y=mass))+
  geom_count(aes(size = ..prop.., group=sxc),alpha=0.5)+
  stat_summary(aes(x=as.numeric(crosstype)),fun.y=mean,geom="line")+ facet_grid(~species)
intplot + aes(group=species, color=species)
```

## Random effects

### Maternal population
```{r randeff_mp}
intplot + aes(group=mompop, color=mompop)
```

### Maternal plant
```{r randeff_mpid}
intplot + aes(group=mompid, color=mompop)
```

### Paternal population
```{r randeff_dp}
intplot + aes(group=dadpop, color=dadpop)
```


## Run models on subsets
Run many generalized linear models on subsets of the data defined by **crosstype | mompid** to see if effects estimates are consistent within maternal plants.

Most maternal plant subsets agree, but some are problematic outliers. These plants can be picked out visually from the random effects interaction plot above, the estimated parameters of each subset model, and the QQ plot of the estimated parameters:

```{r lmlist, results=FALSE, warning=FALSE}
 #had to get rid of species or mompid since mompid is nested inside species. dadpop also works
glm.lis <- lmList(log10(mass)~crosstype|mompid,data=ib, family="gaussian")
plot.lmList(glm.lis,scale=list(x=list(relation="free")))
```

```{r qqmath, results=FALSE, warning=FALSE}
qqmath.lmList(glm.lis)#
```


# Models
We constructed the following models with the package *glmmADMB*. They all have the same fixed effects, **species x crosstype**, and response variable, **log10(mass)**

* X  = standard GL(M)M

| Distribution, Random Effects: | None | Maternal plant | Maternal population |
|-------------------------------|------|----------------|---------------------|
| normal (norm) | X | X | X |

```{r models, echo=T, cache=T}
#Normal (Gaussian) distribution, identity link
ib <- ib[!is.na(ib$collect.date),]
sc.norm.l          <- lm(log10(mass)~species*crosstype+collect.date, data=ib)
sc.mix.mompid.l           <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompid), data=ib)
sc.nd.mix.momdadpid.l           <- lmer(log10(mass)~species*crosstype + (1|mompid)+ (1|dadpid), data=ib)
sc.mix.mompop.l    <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompop), data=ib)
sc.mix.momdadpid.l           <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompid) + (1|dadpid), data=ib)
sc.mix.momdadpop.l    <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompop) + (1|dadpop), data=ib)
```

# Model comparison
## AIC
We will use the Aikake Information Criterion to pick the model the best fits the data, penalized by the number of parameters. Differences of 2 units are significant.


```{r aic}
sc.names <- c("sc.norm.l","sc.mix.mompid.l","sc.mix.mompop.l","sc.mix.momdadpid.l","sc.mix.momdadpop.l")
sc.list <- sapply(sc.names, get, USE.NAMES=T)
sc.AIC <- ICtab(sc.list,mnames=sc.names,type="AIC", base=T, delta=F) # for AICc, nobs=nobs(sc.list[[1]])
class(sc.AIC)<-"data.frame"
all.names <- c(sc.names)
all.list <- sapply(all.names, get, USE.NAMES=T)
all.AIC <- dfun(rbind(sc.AIC))
all.AIC <- all.AIC[order(all.AIC$dAIC),]
kable(all.AIC, format.arg=list(digits=3))
```

The chosen model is a mixed model with the following components:

* 
     + response: **log10(mass)**
     + fixed effects: **species**, **crosstype**, **species x crosstype**
     + random effect: **mompid**


## Overdispersion

Looking at the normal, fixed effects model, we see that the residuals are not normal:
```{r shapiro}
shapiro.test(sc.norm.l$residuals)#raw residuals!
```

## Coefficients
The coefficients estimated for each model agree qualitatively.


```{r coef, fig.asp=1}
sc.log.names <- sc.names
sc.log <- sapply(sc.log.names, get, USE.NAMES=T)

coefplot2(sc.log, legend.x="topright",legend=T,legend.args=list(cex=0.8, xpd=T, inset=c(-0.1,0)), col.pts=sample(gg_color_hue(length(sc.log.names))), spacing=0.05, lwd.2=2, lwd.1=4, intercept=F)
```

# Inference
We chose the model with nearly the best (lowest) AIC, to carry out inference tests and parameter estimation.

## Description


```{r pick}
mod <- sc.mix.momdadpid.l
print(mod)
```


## Test significance of interaction
By dropping it from the model and performing a likelihood-ratio test, we see that the **species x crosstype** interaction is significant:


```{r testint}
sxc.chisq <- drop1(mod, test="Chisq") #load from file
dfun(sxc.chisq)
```

## Model summary
The model estimated the following parameters, with individual parameter significance determined by the Wald z-test, and fixed effect significance determined by analysis of deviance Wald test.


```{r summ}
summary(mod)
Anova(mod, type=3)
```

## Least square means
The least square means procedure can generate predictor estimates of each type, and give their significance groupings with a post-hoc Tukey test. *S. hookeri*-produced hybrids produce less Vegetative Biomass than either crosses between or within *S. hookeri* populations. The other differences are not significant, but remember that the fixed effect of hybrid (vs. between) was significant (model summary).


```{r lsmeans, fig.asp=1, fig.width=5}
#Count
rg <- ref.grid(mod)
#summary(rg)
sxc.lsm <- lsmeans(rg, ~ crosstype*species)
plot(sxc.lsm)
options(digits=4)
cld.mod <- cld(sxc.lsm, Letters=letters) #tukey letterings
cld.mod$response <- 10 ^ cld.mod$lsmean
cld.mod$uSE <- 10 ^ (cld.mod$lsmean+cld.mod$SE)
cld.mod$lSE <- 10 ^ (cld.mod$lsmean-cld.mod$SE)
cld.mod[rev(order(cld.mod$species, cld.mod$crosstype)),]

H.wb <-  with(cld.mod[cld.mod$species=="hook",], response[crosstype=="within"]/response[crosstype=="between"] - 1)
K.wb <- with(cld.mod[cld.mod$species=="kaal",], response[crosstype=="within"]/response[crosstype=="between"] - 1)
K.H <- with(cld.mod[cld.mod$crosstype=="between",], response[species=="kaal"]/response[species=="hook"] - 1)
maxsp <- ifelse(K.H>0, "kaal","hook")
minsp <- ifelse(K.H<0, "kaal","hook")
maxresp <- with(cld.mod, response[species==maxsp & crosstype=="between"])
minresp <- with(cld.mod, response[species==minsp & crosstype=="between"])
HK.resp <-  with(cld.mod, response[species=="hook" & crosstype=="hybrid"])
KH.resp <-  with(cld.mod, response[species=="kaal" & crosstype=="hybrid"])
HK.int <-   with(cld.mod, ifelse(HK.resp > minresp & HK.resp < maxresp, (HK.resp-minresp)/(maxresp-minresp), 
                                 ifelse(HK.resp < minresp, HK.resp/minresp-1, HK.resp/maxresp-1)))
KH.int <-   with(cld.mod, ifelse(KH.resp > minresp & KH.resp < maxresp, (KH.resp-minresp)/(maxresp-minresp), 
                                 ifelse(KH.resp < minresp, KH.resp/minresp-1, KH.resp/maxresp-1)))

intermed <- (minresp + maxresp) / 2
with(ib, wilcox.test(ib[species=="kaal" & crosstype=="hybrid","mass"], mu=intermed))
with(ib, wilcox.test(ib[species=="hook" & crosstype=="hybrid","mass"], mu=intermed))

round(c(H.wb,K.wb,K.H,HK.int,KH.int),2)

ggplot(as.data.frame(cld.mod), aes(y=response, x=relevel(crosstype, "within"), fill=species)) +
  geom_col(position=position_dodge2()) +
  geom_linerange(aes(ymin=lSE, ymax=uSE), position=position_dodge(0.9)) +
  labs(x="", y="Vegetative biomass (g)",fill="Maternal species") +
  scale_fill_manual(labels = c("S. hookeri  ", "S. kaalae  "), values=c("grey70","grey30")) +
  scale_x_discrete(labels = c("Intrapopulation", "Interpopulation", "Hybrid")) +
  geom_text(aes(label=.group), position=position_dodge(0.9), hjust=0, vjust=-1) +
  scale_y_continuous(expand = expand_scale(add=c(0,0)), breaks = scales::pretty_breaks(n = 5)) +
  theme_classic() + theme(legend.text=element_text(face="italic", size=rel(1)), legend.position="bottom", axis.text = element_text(colour="black", size=rel(1)), text=element_text(size=14), axis.ticks.x = element_blank()) + geom_segment(aes(x=2.5, y=intermed, xend=3.5, yend=intermed))
```


