---
title: "Schiedea interfertility"
author: "[John Powers](http://sites.uci.edu/powers/)"
date: "`r Sys.Date()`"#2017-10-19
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: TRUE
  pdf_document:
    toc: yes
resource_files:
- .httr-oauth
- fertiity.csv
- sdata.R
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment="  ", cache=T)
library(arcdiagram)
library(igraph)
library(aod)
library(ggplot2)
library(viridis)
library(ape)


sch <- read.nexus("input/RAxML_1ps_nocp_bipartitionsBranchLabels.sch.short.nobs.nex")
#sch <- read.nexus("input/fig2.nex")
sch$tip.label <- toupper(paste("",substr(sch$tip.label,1,4),sep=""))

species1 <-read.table("input/species.csv",header=TRUE,sep="\t",row.names=2) #load species data
species1$SpName <- factor(species1$SpName)
species1 <- species1[sch$tip.label[-(33:35)],]
```

Does interfertility decrease with phylogenetic distance in *Schiedea*?

##Load data
Data from Weller et al. 2001. F1 hybrids that made greater (green) or less than (red) 50 percent fertile pollen.

```{r data}
fert <- read.table(file="input/fertility/fertility.csv", sep="\t", header=T)
fert$sp1 <- toupper(fert$sp1)
fert$sp2 <- toupper(fert$sp2)
fert$g50 <- fert$g50+1
fspecies <- data.frame(name=levels(factor(c(fert$sp1,fert$sp2))))
g <- graph_from_data_frame(fert, directed=F, vertices=fspecies)
g_edges = get.edgelist(g)
```

##Clustering
```{r cluster}
#clustering by network
adj <- get.adjacency(g, attr="g50")
adj[adj==2] <- 0.5
g.d <- graph_from_adjacency_matrix(adj, weighted=T, mode="undirected")
ceb <- cluster_edge_betweenness(g.d)

#trim phlyo
sch.f <- sch
sch.f$tip.label <- toupper(paste("",substr(sch.f$tip.label,1,4),sep=""))
sch.f<-drop.tip(sch.f,sch.f$tip.label[-match(fspecies$name, sch$tip.label)])

#cophenetic plot
library(phytools)
ceb_tree <- cophylo(as_phylo(ceb), sch.f)
plot(ceb_tree)
```

##Arcplot
```{r arcplot}
#phylo-arcplot
par(mfrow=c(1,2))
par(mar=c(1.2,0,1.2,0))
plot(sch.f)
par(mar=c(0,0,0,0))
arcplot(g_edges, show.labels=F, horizontal=F, ordering=sch.f$tip.label, col.arcs=fert$g50+1)
```

##Fertility vs. phylogenetic distance
```{r pf}
adj <- as.matrix(get.adjacency(g, attr="g50"))
adj[upper.tri(adj, diag=T)]<-NA
adj[adj==0] <- NA
phydist <- cophenetic.phylo(sch.f)
phydist[upper.tri(phydist, diag=T)]<-NA
pf <- na.omit(data.frame(pd=c(phydist), ft=c(adj-1)))
cor(pf$ft, pf$pd, use="complete.obs")
cor.test(pf$ft, pf$pd, use="complete.obs")
```

##Logistic regression
```{r logreg}
#Logistic regression 
model <- glm(ft~pd, family=binomial(link='logit'), data=pf)
summary(model)
confint(model)
wald.test(b = coef(model), Sigma = vcov(model), Terms = 2)
```

##Logistic regression plot
```{r lrplot}
plot <- ggplot(pf, aes(pd, ft)) + stat_smooth(method="glm", method.args=list(family="binomial"), se=T, color="black") + scale_y_continuous(breaks=c(0, 1), labels=c("<50%", ">50%")) + scale_x_continuous() + xlab("Phylogenetic distance") + ylab("Hybrid pollen fertility") +  scale_fill_gradientn(colors=c("white","black")) + theme_bw()

plot + geom_bin2d()
```

```{r data2}
crosses <- read.table("input/fertility/hpol_crosses.csv", sep="\t", header=T, colClasses = list(cross="factor",crosstype="factor"))
str(crosses)
pol <- read.table("input/fertility/hpol.csv", sep="\t", header=T, colClasses = list(cross="factor", id="factor",crosstype="factor"))
pol$vp <- pol$viable/200
pol$mombs <- factor(ifelse(pol$crosstype %in% c("1","2"), "H", ifelse(pol$crosstype=="z", "z", "D")))
pol$dadbs <- factor(ifelse(pol$crosstype %in% c("1","3"), "H", ifelse(pol$crosstype=="z", "z", "D")))
pol$momgrp <- with(pol, factor(paste(momclade,momisle,mombs)))
pol$dadgrp <- with(pol, factor(paste(dadclade,dadisle,dadbs)))
pol$momsp <- crosses$momsp[match(pol$cross, crosses$cross)]
pol$dadsp <- crosses$dadsp[match(pol$cross, crosses$cross)]
str(pol)
heatmap(with(pol, table(momsp,dadsp)), col=viridis(512), scale="none", Colv=NA, Rowv=NA)

#polmeans <- aggregate(vp~cross*momgrp*momclade*momisle*mombs*dadgrp*dadclade*dadisle*dadbs, data=pol, FUN=function(x) c(mean(x), sd(x), length(x)))
polmeans <- aggregate(vp~momsp*dadsp, data=pol, FUN=function(x) c(median(x), quantile(x,0.25), quantile(x,0.75), length(x))) #mean is median!!!
polmeans <- do.call(data.frame, polmeans)
colnames(polmeans)[3:6] <- c("mean", "q25", "q75", "N")
write.csv(polmeans,  "results/hpol_means.csv", row.names=F)
#alllvl <- sort(unique(c(levels(polmeans$momgrp), levels(polmeans$dadgrp))))
#grps <- cbind(alllvl, read.table(text=alllvl, sep=" ", col.names=c("clade","isle","bs")))
species <- unique(c(levels(pol$momsp),levels(pol$dadsp)))
grps <- data.frame(sp=species) #

library(reshape2)
poltbl <- dcast(polmeans, momsp ~ dadsp, value.var="mean", fun.aggregate = mean)
rownames(poltbl) <- poltbl$momsp
poltbl$momsp <- NULL
poltbl2 <- as.matrix(poltbl)
poltbl2[is.nan(poltbl2)] <- 0
```


```{r plot2}
ggplot(pol, aes(x=momsp, color=dadsp, y=vp)) + geom_boxplot()
ggplot(pol, aes(x=momclade, color=dadclade, y=vp)) + geom_boxplot() + geom_point()
ggplot(pol, aes(x=momisle, color=dadisle, y=vp)) + geom_boxplot() + geom_point()
ggplot(pol, aes(x=momclade, color=dadclade, y=vp)) + geom_boxplot() + geom_point()
ggplot(pol, aes(x=reorder(cross, viable, FUN=mean), y=vp, color=crosstype)) + geom_boxplot() + geom_point()
ggplot(pol, aes(x=crosstype, y=vp, color=dadclade, fill=momclade)) + geom_boxplot()

heatmap(poltbl2, col=viridis(512), scale="none", Rowv = NA, Colv=NA)

#ggplot(polmeans, aes(x=factor(momsp, levels=alllvl), y=factor(dadgrp, levels=alllvl), fill=mean)) + geom_tile() + scale_x_discrete(drop=F) +  scale_y_discrete(drop=F) + scale_fill_viridis_c()
#ggplot(polmeans, aes(x=momclade, y=dadclade, fill=mean)) + geom_tile() + scale_fill_viridis_c()
ggplot(polmeans, aes(x=momsp, y=dadsp, fill=mean)) + geom_tile() + scale_fill_viridis_c()

hist(pol$viable, breaks=40)
```


```{r networks2}
colfunc <- viridisLite::plasma

g <- graph_from_data_frame(polmeans[,c("momsp","dadsp","mean","q25","q75","N")], directed=T, vertices=grps)
lgw <-  layout_in_circle(g)
#plot.igraph(g, layout=layout_in_circle(g, order=order(grps$clade)), edge.color=colfunc(100)[round(E(g)$mean*100)],  vertex.color=rainbow(10)[grps$clade], vertex.label.color=rainbow(10)[as.integer(grps$isle)+6], vertex.label.cex=2)
plot.igraph(g, layout=layout_in_circle(g))
  
gu <- as.undirected(g, mode="collapse", edge.attr.comb = "mean")
adj <- get.adjacency(gu, attr="mean")
g.w <- graph_from_adjacency_matrix(adj, weighted=T, mode="undirected")
#plot(g.w, layout_with_fr(g.w), edge.color=colfunc(100)[round(E(g)$mean*100)], edge.width=2,   vertex.color=rainbow(10)[grps$clade], vertex.label.color=rainbow(10)[as.integer(grps$isle)+6], vertex.label.cex=2)
plot(g.w, layout_with_fr(g.w), edge.color=colfunc(100)[round(E(g)$mean*100)], edge.width=2)

#dendPlot(cle, mode="phylo", use.edge.length = T, main="leading eigen")
plot(as.dendrogram(cluster_leading_eigen(gu)))

g_edges = get.edgelist(g)
arcplot(g_edges, show.labels=T, horizontal=F, col.arcs=colfunc(100)[round(polmeans$mean*100)], sorted=T)

```


##Clustering
```{r cluster2}
ceb <- cluster_leading_eigen(gu)
#trim phlyo
sch.f <- sch
sch.f$tip.label <- tolower(paste("",substr(sch.f$tip.label,1,4),sep=""))
sch.f<-drop.tip(sch.f,sch.f$tip.label[-match(species, sch.f$tip.label)])
species1f <- species1
rownames(species1f) <- tolower(rownames(species1f))
species1f <- species1f[sch.f$tip.label,]
#BScols <- setNames(c("black", "navy", "lightblue", "white", "white", "white"),1:6)
BScols <- setNames(RColorBrewer::brewer.pal(6,"Set2")[c(6,4,2,5,1,3)],1:6)
#intraspecies <- setNames(polmeans$mean[match(paste(sch.f$tip.label,sch.f$tip.label), paste(polmeans$momsp,polmeans$dadsp))], sch.f$tip.label)
intraspecies <- setNames(polmeans$mean[match(paste(names(V(g)),names(V(g))), paste(polmeans$momsp,polmeans$dadsp))], names(V(g)))


#cophenetic plot
ceb_tree <- cophylo(as_phylo(ceb), sch.f)
plot(ceb_tree)
```

##Arcplot
```{r arcplot2}
#phylo-arcplot
png("plots/arcplot.png", height=1500, width=3000, pointsize=56)
par(mfrow=c(1,3) , bg="grey30")
par(mar=c(1.5,0,1.5,0))
plot(sch.f, align.tip.label=T, tip.color=BScols[as.integer(species1f$Breeding.System)], edge.width=8, edge.color="grey80", cex=1.5, lwd=8)
add.scale.bar(y=0.002, x=0.002, lcol="white",col="grey80")
par(mar=c(0,0,0,0))
arcplot(g_edges, show.labels=F, horizontal=F, ordering=sch.f$tip.label, col.arcs=plasma(512, direction=-1)[round(polmeans$mean*512)], sorted=T, show.nodes=T, lwd.arcs=6, col.nodes = plasma(512, direction=-1)[round(intraspecies*512)], pch.nodes = 19, cex.nodes=1)
plot.new()
legend(x=-0.05,y=1, legend = rev(levels(species1f$Breeding.System)), col = rev(BScols), pch = 19, cex=1.2, pt.cex=2.5, bty="n", xpd=T, text.col= rev(BScols))
library(plotrix)
color.legend(0,0.55,1,0.65,legend=seq(0,1,0.2),rect.col=plasma(512, direction=-1),cex=1, col="grey80", align="rt")
text(0.2, 0.7, "Pollen viability", col="grey80", cex=1.5)
dev.off()
```

##Fertility vs. phylogenetic distance
```{r pf2}
phydist <- cophenetic.phylo(sch.f)
#phydist[upper.tri(phydist, diag=T)]<-NA

adj <- as.matrix(as_adj(g, attr="mean"))
#diag(adj) <- NA
#adj[upper.tri(adj, diag=T)]<-NA
adj[adj==0] <- NA
adj <- adj[sch.f$tip.label,sch.f$tip.label]

adj.N <- as.matrix(as_adj(g, attr="N"))
adj.N <- adj.N[sch.f$tip.label,sch.f$tip.label]

adj.q25 <- as.matrix(as_adj(g, attr="q25"))
adj.q25 <- adj.q25[sch.f$tip.label,sch.f$tip.label]
adj.q75 <- as.matrix(as_adj(g, attr="q75"))
adj.q75 <- adj.q75[sch.f$tip.label,sch.f$tip.label]

comb <-function(x,y) {mapply(function(q,r) paste(sort(c(q,r)), collapse="-"), x, y)}
labels <- outer(sch.f$tip.label, sch.f$tip.label, comb)

pf <- na.omit(data.frame(pd=c(phydist), ft=c(adj), N=c(adj.N), q25=c(adj.q25), q75=c(adj.q75),crossname= c(outer(sch.f$tip.label, sch.f$tip.label, paste, sep=" x ")), crossname.sort=c(labels)))
pf$q75[pf$q75==1]<- NA # safe for logit transform
cor(pf$ft, pf$pd, use="complete.obs")
cor.test(pf$ft, pf$pd, use="complete.obs")


pdft <- ggplot(pf, aes(x=pd, y=ft, weight=200*N)) + 
    geom_line(aes(group=crossname.sort), col="grey70", size=2, alpha=0.5) + 
  theme_grey() + 
  xlab("Phylogenetic distance (mutations/site)") + 
  ylab("Pollen viability")

pdft + geom_point() + geom_linerange(aes(ymin=q25, ymax=q75)) + ylim(c(0,1))

pdft + scale_y_continuous(trans="logit", breaks=c(0.01,seq(0,1,by=0.05), 0.995)) + geom_text(aes(label=crossname), size=3) # trans_breaks(function(p) { log(p/(1-p)) }, function(x) { 1/(1+exp(-x)) },  n=10)

pdft  + geom_smooth(method="glm", method.args=list(family="binomial"), color="grey70", fill="grey80") + geom_text(aes(label=crossname, color=N)) + coord_trans(y="logit") + scale_y_continuous(breaks=c(0.01,seq(0,1,by=0.05), 0.994)) + scale_color_gradient(trans="sqrt", low="darkblue", high="green2")
#scale_y_continuous(breaks=seq(0,1, by=0.05), limits=c(0,1), expand=c(0,0))

ggsave("pdft.png", height=9, width=9)
```

##Arcplot
```{r arcplot3}
#phylo-arcplot
#TODO: arrange arcs front-to-back by phydist?
library(gridBase)
library(grid)

png("plots/arcplot.png", height=1500, width=4000, pointsize=20, res=300)
par(mfrow=c(1,4) , bg="grey30")
par(mar=c(1,0,1,0))
plot(sch.f, align.tip.label=T, tip.color=BScols[as.integer(species1f$Breeding.System)], edge.width=3, edge.color="grey80", cex=1, lwd=4, label.offset=0.0001)
add.scale.bar(y=0.6, x=0, lcol="grey80",col="grey80", cex=1, lwd=3)
par(mar=c(0,0,0,0))
arcplot(g_edges, show.labels=F, horizontal=F, ordering=sch.f$tip.label, col.arcs=plasma(512, direction=-1)[round(polmeans$mean*512)], sorted=T, show.nodes=T, lwd.arcs=3, col.nodes = plasma(512, direction=-1)[round(intraspecies[order(names(intraspecies))]*512)], pch.nodes = 19, cex.nodes=1)
plot.new()
legend(x=-0.05,y=1, legend = rev(levels(species1f$Breeding.System)), col = rev(BScols), pch = NA, cex=1, pt.cex=2.5, bty="n", xpd=T, text.col= rev(BScols))
library(plotrix)
color.legend(0.1,0.05,0.2,0.55,legend=seq(0,1,0.2),rect.col=plasma(511, direction=-1),cex=0.45, col="grey80", align="rt", gradient="y")
text(0.26, 0.6, "Pollen viability", col="grey80", cex=1)
plot.new()    
vps <- baseViewports()
pushViewport(vps$figure) 
vp1 <-plotViewport(c(0,0,0,0)) 
print(pdft  + 
        geom_smooth(method="glm", method.args=list(family="binomial"), color="grey70", fill="grey80") +
        geom_linerange(aes(ymin=q25, ymax=q75)) +
        geom_point(aes(color=ft), size=2) + 
        coord_trans(y="logit") + 
        scale_y_continuous(breaks=c(0.01,seq(0,1,by=0.1), 0.99)) + 
        scale_color_viridis_c(option="plasma", direction=-1) + 
        theme_classic() + 
        theme(plot.background = element_rect(color = "grey30", fill = "grey30"), panel.background = element_rect(fill = "grey30", color  =  NA),  legend.background = element_rect(color = NA, fill = "grey30"), axis.text = element_text(color = "grey80"), axis.title = element_text(color = "grey80")) + 
        guides(color=F), vp = vp1)
dev.off()
```

##Binomial regression on means
```{r logreg2}
model <- glm(ft~pd, family="quasibinomial", weights=N*200, data=pf)
summary(model)
plot(model)

library(glmmADMB)
betabin.admb <- glmmadmb(round(cbind(ft*N*200,(1-ft)*N*200))~pd, data=pf, family="betabinomial")
summary(betabin.admb)
```


##Binomial regression on raw data 
```{r logreg2}
phydist.m <- melt(phydist)
polmeans$pd <- phydist.m$value[match(paste(polmeans$momsp,polmeans$dadsp), paste(phydist.m$Var1,phydist.m$Var2))]
pol$pd <- phydist.m$value[match(paste(pol$momsp,pol$dadsp), paste(phydist.m$Var1,phydist.m$Var2))]
pol$tot <- 200
pol$inviable <- pol$tot-pol$viable

ggplot(pol, aes(y=vp,x=pd)) + geom_point()

model <- glm(cbind(viable, inviable)~pd, family="quasibinomial", data=pol)
summary(model)
#plot(model)
library(jtools)
effect_plot(model, pred = pd, interval = TRUE, data=pol) + geom_point(data=pol, aes(pd,vp), color="blue")
coef(model)

library(glmmADMB)
plot.glmmadmb <- function(x, ...) {  qplot(x$fitted, x$residuals, xlab="Fitted values", ylab="Residuals", ...) + geom_smooth()}

betabin.admb <- glmmadmb(cbind(viable,inviable)~pd, data=pol, family="betabinomial")
summary(betabin.admb)
plot(betabin.admb)
plot(fitted(betabin.admb)~pol$pd)
points(polmeans$mean~polmeans$pd, col="red")
coef(betabin.admb)

#betabin.admb.momdadsp <- glmmadmb(cbind(viable,inviable)~pd + (1|momsp) + (1|dadsp), data=pol, family="betabinomial")
load("betabin.admb.momdadsp.Rdata")
summary(betabin.admb.momdadsp)
plot(betabin.admb.momdadsp)
#save(betabin.admb.momdadsp, file="betabin.admb.momdadsp.Rdata")

######Plots############
#Looks like the random effects are overfitting :( GLM does the best
#pol$fitted <- fitted(betabin.admb)
#pol$fitted <- fitted(betabin.admb.momdadsp)
pol$fitted <- fitted(model)
polmeans$fitmean <- aggregate(fitted~momsp*dadsp, data=pol, FUN=mean)$fitted
  ggplot(polmeans, aes(pd, fitmean)) +     geom_smooth(aes(y=mean), method="glm", method.args=list(family="binomial"), color="red", fill="pink") + geom_smooth(method="glm", method.args=list(family="binomial"), color="black") + geom_point() + geom_point(aes(y=mean), color="red") + geom_segment(aes(xend=pd, yend=mean))  + geom_text(aes(y=mean, label=paste(momsp,dadsp,sep=" x ")), color="blue")

####Random effects#####
momsp.ef <- data.frame(sp=rownames(ranef(betabin.admb.momdadsp)[[1]]), ef=as.vector(ranef(betabin.admb.momdadsp)[[1]]), sd=as.vector(ranef(betabin.admb.momdadsp, sd=T)[[1]]))
momsp.ef$sp <- reorder(momsp.ef$sp, momsp.ef$ef)
ggplot(momsp.ef, aes(sp,ef)) + geom_point() + geom_linerange(aes(ymin=ef-sd, ymax=ef+sd)) + ggtitle("Maternal random effects") + xlab("Species") + ylab("Effect")

dadsp.ef <- data.frame(sp=rownames(ranef(betabin.admb.momdadsp)[[2]]), ef=as.vector(ranef(betabin.admb.momdadsp)[[2]]), sd=as.vector(ranef(betabin.admb.momdadsp, sd=T)[[2]]))
dadsp.ef$sp <- reorder(dadsp.ef$sp, dadsp.ef$ef)
ggplot(dadsp.ef, aes(sp,ef)) + geom_point() + geom_linerange(aes(ymin=ef-sd, ymax=ef+sd)) + ggtitle("Paternal random effects") + xlab("Species") + ylab("Effect")


```
