---
title: '*Schiedea* pollen tube growth'
author: "[John Powers](http://sites.uci.edu/powers/)"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
editor_options:
  chunk_output_type: console
---

# Purpose
Test for post-pollination, pre-zygotic reproductive barriers between five species pairs of *Schiedea* that have hybridization potential, by measuring pollen tube growth from parental and hybrid crosses at two time points.

Response variables:

* Counts - assume a Poisson distribution for count data 
    - **grains** - the number of pollen grains on the stigma (only measured for some kaahoo crosses)
    - **mid** - the number of pollen tubes counted at the middle of the style
    - **bottom** - the number of pollen tubes counted at the bottom of the style

* Proportions - assume a binomial distribution for binomial data
    - **ratio** - the ratio of pollen tubes at the **bottom** / **middle** of the style, capped at 1 because having more pollen grains at the bottom than the middle of the style is impossible (must result from counting error). For the binomial model, the input is the number of successes (pollen tubes at **bottom**) and failures (min(**bottom**/**mid**) - **bottom**).

Fixed effects:

* **cross** - the maternal plant x the paternal plant
* **hr** - the time elapsed since pollination: 4 hr or 24 hr

Random effects:

* **momid** - maternal genotype, specified by its cross and plant number
* **dadid** - paternal genotype, specified by its cross and plant number
* **vial** - vial that holds the stigmas from one pollination

```{r setup, message=FALSE}
library(knitr)
knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, width=999,
                      fig.path = "figures/", dev = "svglite", dev.args=list(fix_text_size=FALSE))
options(knitr.kable.NA = '')
library(tidyverse)
library(gridExtra)
library(ggsignif) 
library(gridExtra)
library(broom)
library(emmeans) #Estimated marginal means aka least square means.
library(glmmTMB) #generalized linear mixed models with Template Model Builder

#overwrite function to make emmeans and multcomp play nice
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
    multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                        df = df, ...)
}

signif.num <- function(x) {
   cut(x, breaks = c(0, 0.001, 0.01, 0.05, 1), 
           labels = c("***", "**", "*", " "))
}
```

# Read data

```{r read}
#long format: salmen (1 stigma/vial), lydsar (2 stigmas/vial), kaahoo (3 stigmas/vial without letter)
#wide format: memkau (3-5 stigmas/vial), speste (6 stigmas/vial), kaahoo_extra(3-5 stigmas/vial)
pops <- list(sali="842", menz = c("849", "950"), 
             lydg="6682", sarm="896", 
             sper= c("120289","120294","120296"), stel = c("14651","14765"),
             memb = c("mem","864"), kaua = "kaua",
               kaal = c("892", "3587"), hook = c("794", "866", "879","899"))

pair_labels <- c(salmen = "S. salicaria | S. menziesii",
                 lydsar = "S. lydgatei | S. sarmentosa",
                 speste = "S. spergulina | S. stellariodes",
                 memkau = "S. membranacea | S. kauaiensis",
                 kaahoo = "S. kaalae | S. hookeri")

abbrev <- list(salmen = c(S="Sl",M="Mn"), 
               lydsar = c(L="Ly",S="Sr"), 
               speste = c(Sp="Sp",St="St"), 
               memkau = c(M="Mb",K="Ki"),
               kaahoo = c(K="Kl",H="Ho"))

species_names <- c("Mn"="menziesii", "Sl"="salicaria",
              "Ly"="lydgatei", "Sr" ="sarmentosa",
              "St"="stellarioides","Sp"="spergulina",
              "Mb"="membranacea","Ki"="kauaiensis",
              "Kl"="kaalae", "Ho"="hookeri")

cross_levels <- map(abbrev, ~ c(paste0(.x[1],"x",.x[1]), paste0(.x[1],"x",.x[2]), paste0(.x[2],"x",.x[2]), paste0(.x[2],"x",.x[1])))

tubesl <- list()             
# kaalae - hookeri
tubesl[["kaahoo"]] <- read_csv("data/pollentubes_kaahoo_extra.csv") %>% 
  pivot_longer(starts_with(c("mid","bottom")), names_to=c("location","stigma"), names_sep="_", values_to="n_tubes") %>% 
  pivot_wider(names_from="location", values_from="n_tubes") %>% 
  drop_na(mid, bottom) %>% #one entry has mid but no bottom count
  mutate(momcross = toupper(str_sub(momsp,1,1)), dadcross = toupper(str_sub(dadsp,1,1)),
         observer="SGW") %>% 
  bind_rows(read_csv("data/pollentubes_kaahoo.csv") %>% #edited from pollentubes.csv
              separate(cross, into=c("momcross","dadcross"), sep="x") %>% 
              mutate(observer="SZ",
                     vial = as.numeric(str_remove(vial, "[ABCD]")))) %>% 
  mutate(momid = paste(momcross, momid), dadid = paste(dadcross, dadid))

# lydgatei- - sarmentosa
tubesl[["lydsar"]] <- read_csv("data/pollentubes_lydsar.csv") %>% 
  filter(keep == 1, hr != 2) %>% select(-keep) %>%  #exclude the resamples of pollinations that gave no tubes, and counts at 2 hrs
  mutate(mompop = as.character(mompop), dadpop=as.character(dadpop),
         momid = paste(mompop, momplant), dadid = paste(dadpop, dadplant),
         momcross = ifelse(mompop %in% pops[["sarm"]], "S", "L"), dadcross = ifelse(dadpop %in% pops[["sarm"]], "S", "L"))

#salicaria - menziesii
tubesl[["salmen"]] <- read_csv("data/pollentubes_salmen.csv") %>% 
  select(-cross) %>% 
  mutate(mompop = str_sub(momid, 1, 3), dadpop = str_sub(dadid, 1, 3),
         momcross = ifelse(mompop %in% pops[["sali"]], "S", "M"), dadcross = ifelse(dadpop %in% pops[["sali"]], "S", "M"))

# membranacea - kauaiensis
tubesl[["memkau"]] <- read_csv("data/pollentubes_memkau.csv") %>% mutate(across(ends_with("sp"), as.factor)) %>% 
  pivot_longer(starts_with(c("mid","bottom")), names_to=c("location","stigma"), names_sep="_", values_to="n_tubes") %>% 
  pivot_wider(names_from="location", values_from="n_tubes") %>% 
  mutate(status = fct_explicit_na(status, "OK"),
         mid =    ifelse(status=="no pollen on stigmas", NA, mid), #set to NA to drop, 0 to keep
         bottom = ifelse(status=="no pollen on stigmas", NA, bottom)) %>% #see SGW email 2022-05-09, NOTES tab
  drop_na(mid) %>% 
  mutate(momcross = ifelse(momsp %in% pops[["memb"]], "M", "K"),  dadcross = ifelse(dadsp %in% pops[["memb"]], "M", "K"),
         momid = paste(momcross, momplant), dadid = paste(dadcross, dadplant),)

#spergulina - stellarioides
tubesl[["speste"]] <- read_csv("data/pollentubes_speste.csv") %>% mutate(across(ends_with("pop"), as.factor)) %>% 
  pivot_longer(starts_with(c("mid","bottom")), names_to=c("location","stigma"), names_sep="_", values_to="n_tubes") %>% 
  pivot_wider(names_from="location", values_from="n_tubes") %>% drop_na(mid) %>% 
  mutate(momid = paste(mompop, momplant), dadid = paste(dadpop, dadplant),
         momcross = ifelse(mompop %in% pops[["sper"]], "Sp", "St"), dadcross = ifelse(dadpop %in% pops[["sper"]], "Sp", "St"))

tubes <- bind_rows(tubesl, .id = "pair") %>% 
  mutate(hr = factor(paste0(hr,"hr"), levels=c("4hr","24hr")),
         ratioraw = bottom/mid,
         adjmid = ifelse(bottom > mid, bottom, mid),
         ratio = bottom/adjmid,
         justmid = mid - bottom,
         vial = as.character(vial), #for random effect of pollination
         momcross = map2_chr(pair, momcross, ~ abbrev[[.x]][[.y]]),
         dadcross = map2_chr(pair, dadcross, ~ abbrev[[.x]][[.y]]),
         cross = factor(paste0(momcross,"x",dadcross), levels=unlist(cross_levels)),
         crosshr = factor(paste0(cross, hr), levels = paste(rep(levels(cross),each=2), levels(hr), sep="")),
         pair = factor(pair, levels=names(pair_labels))) %>% arrange(pair)
str(tubes)
```

# Inventory of crosses

```{r pollentube_inventory}
kable(with(tubes, table(cross, hr)), caption="Number of styles of each cross type")
kable(with(tubes, table(pair, hr)), caption="Number of styles of each species pair")
kable(tubes %>% count(pair, cross, vial) %>% with(table(pair, n)), caption = "Number of styles counted per vial")
```

# Models

```{r models}
pair_contrasts <- function(mod) {
  lv <- levels(mod$frame$crosshr)
  crosshr.x <- set_names(rep(1:4, each=2) + rep(c(-0.25, 0.25), 4), lv)
  tube.contrasts <- list(c(5,6), c(7,8), c(5,7), c(6,8),
                         c(3,4), c(1,2), c(3,1), c(4,2), 
                         c(5,1), c(6,2)) %>% 
    map_chr(~paste(lv[.x[1]], "-", lv[.x[2]], "== 0"))
  
  multcomp::glht(mod, linfct = multcomp::mcp(crosshr=tube.contrasts)) %>%
  multcomp:::summary.glht(test=multcomp::adjusted(type="fdr")) %>% tidy() %>% 
  separate(contrast, c("lvl1", "lvl2"), sep = " - ", extra="merge") %>%
  mutate(signif = adj.p.value < 0.05, 
         stars = signif.num(adj.p.value), 
         height =  c(rep(c(1,1,2,3),2),4,5), 
         x1 = crosshr.x[lvl1], 
         x2 = crosshr.x[lvl2])
}

pair_emm <- function(mod) {
  emm <- ref_grid(mod) %>% emmeans(~ cross*hr) %>% summary()
  if(mod$modelInfo$family$link == "logit") {
    emm %>% mutate(response = plogis(emmean), uSE = plogis(emmean+SE), lSE = plogis(emmean-SE))
  } else if(mod$modelInfo$family$link == "log") {
    emm %>% mutate(response = exp(emmean), uSE = exp(emmean+SE), lSE = exp(emmean-SE))
  }
}

tubes.mod <- tubes %>% group_by(pair) %>% nest() %>% 
  mutate(mid_mod    = map(data, ~ glmmTMB(mid~cross*hr + (1|momid) + (1|dadid) + (1|vial), data=.x, family="poisson")),
         mid_mod2   = map(data, ~ glmmTMB(mid~crosshr +  (1|momid) + (1|dadid) + (1|vial), data=.x, family="poisson")),
         mid_emm    = map(mid_mod, pair_emm),
         mid_cont   = map(mid_mod2, pair_contrasts),
         bottom_mod = map(data, ~ glmmTMB(bottom~cross*hr + (1|momid) + (1|dadid) + (1|vial), data=.x, family="poisson")),
         bottom_mod2= map(data, ~ glmmTMB(bottom~crosshr  + (1|momid) + (1|dadid) + (1|vial), data=.x, family="poisson")),
         bottom_emm = map(bottom_mod, pair_emm),
         bottom_cont= map(bottom_mod2, pair_contrasts),
         ratio_mod  = map(data, ~ glmmTMB(cbind(bottom, adjmid-bottom)~cross*hr + (1|momid) + (1|dadid) + (1|vial), 
                                          data=.x, family="binomial")),
         ratio_mod2 = map(data, ~ glmmTMB(cbind(bottom, adjmid-bottom)~crosshr  + (1|momid) + (1|dadid) + (1|vial), 
                                          data=.x, family="binomial")),
         ratio_emm  = map(ratio_mod, pair_emm),
         ratio_cont = map(ratio_mod2, pair_contrasts))
```

```{r pollentubes, fig.width=19, fig.height=6}
mytheme <- theme(legend.text=element_text(size=rel(1)), legend.position="right", 
                 axis.text = element_text(colour="black", size=rel(1)), text=element_text(size=14), 
                 axis.ticks.x = element_blank(), plot.title = element_text(face="italic", size=14))
timefills <-   scale_fill_brewer(palette=3, type="qual", labels=c("4 hr","24 hr"))

var_labels <- c(mid = "Pollen tubes at middle of style",
                bottom = "Pollen tubes at bottom of style",
                ratio = "Ratio of pollen tubes at bottom : middle of style",
                justmid = "Difference of pollen tubes at middle and bottom of style")

plot_emm <- function(pair, emm, cont, data, yvar) {
  ggplot(emm, aes(y=response, x=cross, fill=hr)) +
    geom_violin(data=data, aes_string(y=yvar),position=position_dodge(), 
                draw_quantiles=c(0.25,0.5,0.75), linewidth=0.5, bw=ifelse(yvar=="ratio",0.1,1)) +
    geom_pointrange(aes(ymin=lSE, ymax=uSE), position=position_dodge2(0.9)) +
    geom_signif(data=emm, y_position=cont$height/c(mid=1, bottom=1, ratio=15)[yvar]+c(mid=19, bottom=19, ratio=1)[yvar], 
                xmin = cont$x1, xmax = cont$x2, annotations=replace_na(as.character(cont$stars), "NA"), vjust=0.5, tip_length=0.01) +
    labs(x="", y=var_labels[yvar], title=pair_labels[pair]) +
    scale_y_continuous(limits=c(0,c(mid=24, bottom=24, ratio=1.4)[yvar]), 
                       expand = expansion(add=c(0,c(mid=0.5, bottom=0.5, ratio=0.05)[yvar])), 
                       breaks = seq(0,1,by=0.25) * c(mid = 20, bottom = 20, ratio = 1)[yvar])+
    scale_x_discrete(labels=~str_replace(.x, "x"," x ")) + 
    theme_classic() + mytheme + timefills + guides(fill="none")
}

grid.arrange(grobs=pmap(select(tubes.mod, pair, emm=mid_emm, cont=mid_cont, data), 
                        plot_emm, yvar="mid"), nrow=1)
grid.arrange(grobs=pmap(select(tubes.mod, pair, emm=bottom_emm, cont=bottom_cont, data),
                        plot_emm, yvar="bottom"), nrow=1)
grid.arrange(grobs=pmap(select(tubes.mod, pair, emm=ratio_emm, cont=ratio_cont, data),
                        plot_emm, yvar="ratio"), nrow=1)
```

```{r pollentubes_all, fig.width=17, fig.height=6}
plot_all <- function(yvar) {
ggplot(unnest(select(tubes.mod, paste0(yvar,"_emm")), paste0(yvar,"_emm"))) +
  facet_wrap(vars(pair), scales="free_x", nrow=1, labeller = as_labeller(pair_labels)) +
  geom_violin(data=unnest(select(tubes.mod, data), data),  aes(y=.data[[yvar]], x=cross, fill=hr),position=position_dodge(), 
              draw_quantiles=c(0.25,0.5,0.75), linewidth=0.5, bw=ifelse(yvar=="ratio",0.1,1)) +
  geom_pointrange(aes(y=response, x=cross, fill=hr, ymin=lSE, ymax=uSE), position=position_dodge2(0.9)) +
  geom_signif(data=unnest(select(tubes.mod, paste0(yvar,"_cont")), paste0(yvar,"_cont")) %>% 
                mutate(n=row_number(), stars=replace_na(as.character(stars), "NA"), #TODO speste contrasts are failing, fitting problem gives NAs here
                       height=height/c(mid=1, bottom=1, ratio=15)[yvar]+c(mid=19, bottom=19, ratio=1)[yvar]),
              aes(y_position=height, xmin = x1, xmax = x2, annotations=stars, group=n), 
              vjust=0.5, tip_length=0.01, manual=T) +
  labs(x="", y=var_labels[yvar]) +
  scale_y_continuous(limits=c(0,c(mid=24, bottom=24, ratio=1.4)[yvar]), 
                     expand = expansion(add=c(0,c(mid=0.5, bottom=0.5, ratio=0.05)[yvar])), 
                     breaks = seq(0,1,by=0.25) * c(mid = 20, bottom = 20, ratio = 1)[yvar])+
  scale_x_discrete(labels=~str_replace(.x, "x"," x ")) + 
  theme_classic() + mytheme + timefills + guides(fill="none")
}
plot_all("mid")
plot_all("bottom")
plot_all("ratio")
```


# Tubes at bottom v. middle of style

```{r pollentubes_diff, fig.width=9, fig.height=7}
ggplot(tubes, aes(x=bottom, y=mid)) + 
  facet_wrap(vars(pair), labeller=as_labeller(pair_labels))+
  labs(x=var_labels["bottom"], y=var_labels["mid"]) +
  geom_abline(slope=1, size=1) +
  geom_point(alpha=0.1) + 
  coord_fixed(xlim=c(0,NA),ylim=c(0,NA))+
  theme_classic()+mytheme+theme(strip.text=element_text(face="italic"))
```

# Phylogenetic distance

```{r phylo_tubes, fig.height=8}
library(ape)
tree <- read.nexus("data/RAxML_1ps_nocp_bipartitionsBranchLabels.sch.short.nobs.nex")

tubes.phylodist <- tubes.mod %>% select(pair, ends_with("_emm")) %>% pivot_longer(ends_with("_emm"), names_to="trait") %>% 
  mutate(trait = str_remove(trait, "_emm")) %>% unnest(value) %>% 
  mutate(momsp = factor(species_names[str_sub(cross, 1,2)], levels=species_names)) %>% group_by(pair, trait, momsp, hr) %>% 
  summarize(trait_diff = response[1]-response[2], 
            dadsp = species_names[str_sub(cross[2], 4,5)], .groups = "drop") %>% 
  mutate(pd = map2_dbl(as.character(momsp), as.character(dadsp), ~cophenetic(tree)[.x,.y]))

ggplot(tubes.phylodist, aes(x=pd, y=trait_diff)) + facet_grid(trait~hr, scales="free_y") +
  geom_point(aes(color=momsp), size=2) + geom_smooth(method="lm", se=F, color="grey") + 
  scale_color_brewer(palette="Paired")+
  labs(x="Phylogenetic distance (mutations/site)", color="Maternal species",
       y="Difference between intra and interspecies cross") + theme_bw() + 
  theme(legend.position = "top")
```

# Package citations

```{r citations}
citation("glmmTMB")
citation("emmeans")
citation("multcomp")
```

