---
title: "*Schiedea* pollen competition"
author: "[John Powers](http://sites.uci.edu/powers/)"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
---

```{r setup, message = FALSE}
library(tidyverse)
library(glmmTMB)
library(broom)
library(emmeans)
library(knitr)
knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, fig.height=8, fig.width=8,
                      fig.path = "figures-comp/", dev = "svglite", dev.args=list(fix_text_size=FALSE))
options(digits=4, knitr.kable.NA = "")
ggplot2::theme_set(theme_minimal())
```

# Inventory

```{r load_data}
#in order of phylogenetic distance, pair order is by style length
sp_pairs <- c("menz"="menziesii", "sal"="salicaria",
              "lydg"="lydgatei", "sarm" ="sarmentosa",
              "stell"="stellarioides","sperg"="spergulina",
              "memb"="membranacea","kaua"="kauaiensis",
              "kaal"="kaalae", "hook"="hookeri")
typenames <- set_names(c("Intraspecific","Competitive","Interspecific"),c("intra","mixed","inter"))
type_pal <- set_names(c("#DC0060", "#86D400", "#F56E00"), names(typenames))
propnames <- set_names(c("Germinated","Survived","Hybrid"), c("prop.germ", "prop.alive", "prop.hybrid"))

germ.pot <- read_tsv("data/Pollen competition - germination.tsv") %>%  
  mutate(dadspm = if_else(type=="mixed" & dadsp1==momsp, dadsp1, dadsp2), #rearrange dad1/dad2 to have dadsp match momsp
         dadidm = if_else(type=="mixed" & dadsp1==momsp, dadid1, dadid2),
         dadsp  = if_else(type=="mixed" & dadsp1==momsp, dadsp2, dadsp1),
         dadid  = if_else(type=="mixed" & dadsp1==momsp, dadid2, dadid1), .after=momid) %>% 
  select(-starts_with("germinated_n"), -transplanted_n, -matches("dad(sp|id)[12]")) %>% 
  mutate(across(matches("sp"), ~ recode(.x, !!!sp_pairs)),
         sxc = paste(momsp, type, sep="."),
         across(where(is.character), factor), 
         type=factor(type, levels= names(typenames)), 
         across(matches("sp"), ~ factor(.x, levels = sp_pairs)),
         germinated = ifelse(germinated > planted, planted, germinated))

germ <- germ.pot %>% group_by(date_planted, type, momsp, sxc, momid, dadspm, dadidm, dadsp, dadid) %>% 
  summarize(planted = sum(planted), germinated = sum(germinated), .groups="drop") %>% # add together all the pots of one cross
  mutate(prop.germ = germinated/planted) # proportion of planted seeds that germinated 

surv <- read_tsv("data/Pollen competition - survival.tsv") %>% 
  mutate(dadspm = if_else(type=="mixed" & dadsp1==momsp, dadsp1, dadsp2), #rearrange dad1/dad2 to have dadsp match momsp
         dadidm = if_else(type=="mixed" & dadsp1==momsp, dadid1, dadid2),
         dadsp  = if_else(type=="mixed" & dadsp1==momsp, dadsp2, dadsp1),
         dadid  = if_else(type=="mixed" & dadsp1==momsp, dadid2, dadid1), .after=momid) %>% 
  select(-matches("dad(sp|id)[12]")) %>% drop_na(dead) %>% #one row excluded
  mutate(sxc = paste(momsp, type, sep="."),
         across(where(is.character), factor), 
         type=factor(type, levels= names(typenames)),
         across(matches("sp"), ~ factor(.x, levels = sp_pairs)),
         across(all_of(c("sameasmom", "hybrid")), ~replace_na(.x,0)), 
         alive = sameasmom + hybrid,
         total = alive + dead, 
         prop.hybrid = hybrid / alive,
         prop.alive = alive / total)

surv %>% count(momsp, type, dadsp, name="crosses_surv") %>% 
  full_join(germ %>% count(momsp, type, dadsp, name="crosses_germ")) %>% 
  full_join(germ.pot %>% count(momsp, type, dadsp, name = "pots_germ")) %>%
  mutate(pots_per_cross = pots_germ/crosses_germ) %>% 
  kable(caption="Number of crosses between individual plants used for germination and survival,
        and how many pots the seeds filled")

styles <- read_tsv("data/Pollen competition - style_length.tsv") %>% select(-starts_with("px")) %>%
  left_join(tibble(momsp=sp_pairs, pair = factor(rep(1:5, each=2)))) %>% 
  mutate(momsp = factor(momsp, level=sp_pairs)) %>% 
  group_by(pair) %>% mutate(style_diff_mm = style_length_mm-min(style_length_mm))
```

Note that the crosses_surv disagrees with slide 10 for stellarioides and spergulina. 

# Scoring dates

The date scored is not always available or accurate, do not use for analysis.

```{r dates}
germ.pot %>%  ggplot(aes(x=date_planted, y=date_scored, color=momsp)) + 
  geom_abline(slope=1, intercept=0) + geom_abline(slope=1, intercept=90) + geom_point() + geom_rug() +
  scale_color_brewer(palette="Paired") + labs(title="Germination scoring") #TODO check mis-entered dates

germ.pot %>% filter(abs(date_scored-date_planted)<=90) %>% ggplot(aes(x=date_scored - date_planted, fill=type)) +
  facet_wrap(vars(momsp)) +
  geom_histogram(binwidth=1) + scale_x_continuous(limits=c(0,NA)) + 
  labs(title="Germination scoring", subtitle = "plot restricted to 90 days")

surv %>% ggplot(aes(x=date_planted, y=date_scored, color=momsp)) + geom_point() + geom_rug() + 
   scale_color_brewer(palette="Paired") + labs(title="Survival scoring")
```

# Models

Expand this code to see the model specifications. The models have the following settings:

* response variable: proportion seeds that germinated, proportion seedlings that survived, proportion of live seedlings that are hybrids
* sxc: the combined maternal plant species and crosstype (within species, between species, mixed pollen)
* momid: random effects of the maternal plant genotype 
* family: betabinomial (for binary variables)

```{r models}
comp.mod <- list(
  prop.germ =    glmmTMB(prop.germ   ~ sxc + (1|momid), family="betabinomial", data=germ, weights = planted),
  prop.alive =   glmmTMB(prop.alive  ~ sxc + (1|momid), family="betabinomial", data=surv, weights = total),
  prop.hybrid =  glmmTMB(prop.hybrid ~ sxc + (1|momid), family="betabinomial", data=filter(surv, type=="mixed"), weights = alive))
```

## Inference 

P-values for the overall effect of species and crosstype, or collection date, on the trait.

```{r inference}
comp.test <- map_dfr(comp.mod, ~car::Anova(.) %>% tidy(), .id="trait") 
comp.test %>%  mutate(term=recode(term, sxc = "species and crosstype"),
         p.value=format(p.value,digits=2)) %>% 
  kable(digits=1, caption="ANOVA of GLMM (Type III Wald chisquare tests)")
  
comp.emm <-  map_dfr(comp.mod, ~emmeans(., ~sxc) %>% summary(type="response") %>%  #average over collect date
                             tidy %>% rename(estimate=prob), .id="trait") %>% 
  separate(sxc, into = c("momsp","type")) %>% 
  mutate(momsp=factor(momsp, levels=sp_pairs), 
         type=factor(type, levels= names(typenames)),
         trait=factor(trait, levels=names(propnames)))

#doesn't work, compares all to the first level (hookeri)
#hybrid.half <- multcomp::glht(comp.mod$prop.hybrid, multcomp::mcp(sxc = paste0(sp_pairs[-7],".mixed = 0")))
#compare to null of 50% hybrids, since pollen mixtures were 50-50
hybrid.test <- surv %>% filter(type=="mixed") %>% 
  group_by(momsp, type, dadsp) %>% 
  summarize(total=sum(total), alive=sum(alive), hybrid=sum(hybrid), crosses=n(), .groups="drop") %>% 
  mutate(total_per_c = total/crosses, alive_per_c = alive / crosses,
         prop.hybrid = hybrid/alive,
         p = map2_dbl(hybrid, alive, ~prop.test(x=.x, n=.y, p = 0.5)$p.value),
         p.adj = p.adjust(p, "bonferroni"),
         stars = cut(p.adj, breaks = c(-Inf, 0.01, 0.05, 0.10, Inf), 
                labels = c("***", "**", "*", "n.s."), right = FALSE))
```

# Plots

Boxplots of the germination and survival proportions in each plant x plant cross. Germination proportions are calculated across all pots of a given plant x plant cross. Points show estimated marginal mean and 95% CI.

```{r rawdata}
sp_theme <- theme(legend.position = "top", 
                  panel.grid.major.x=element_blank(), panel.grid.minor.y=element_blank(), 
                  axis.text.x = element_text(angle=20), axis.text = element_text(color="black"))
alternate <- geom_tile(data=tibble(momsp=sp_pairs[seq(1,10,by=2)]), aes(x=momsp, y=0.5, height=1, width=1),  alpha=0.15, fill="grey", inherit.aes = F) 

ggplot(germ, aes(x=momsp, fill=type, y=prop.germ))  + 
  geom_boxplot(position = position_dodge(width=0.9), color="grey50", outlier.size=0.5) + alternate+ 
  geom_pointrange(data=filter(comp.emm, trait=="prop.germ"), aes(y=estimate, ymax=conf.high, ymin=conf.low),
                  position = position_dodge(width=0.9), size=0.3, show.legend=F) + 
  scale_fill_manual(values=type_pal, labels=typenames) +  
  labs(x="Maternal species", y="Germination", fill="Pollination") + sp_theme

ggplot(surv, aes(x=momsp, fill=type, y=prop.alive))  + 
  geom_boxplot(position = position_dodge(width=0.9), color="grey50", outlier.size=0.5) + alternate+ 
  geom_pointrange(data=filter(comp.emm, trait=="prop.alive"), aes(y=estimate, ymax=conf.high, ymin=conf.low),
                  position = position_dodge(width=0.9), size=0.3, show.legend=F) + 
  scale_fill_manual(values=type_pal, labels=typenames) +  
  labs(x="Maternal species", y="Survival", fill="Pollination") + sp_theme

ggplot(filter(surv, type=="mixed"), aes(x=momsp, y=prop.hybrid, fill=momsp)) + 
  geom_hline(yintercept=0.5)+ geom_violin(scale="width", draw_quantiles = c(0.25,0.5,0.75))+ 
  geom_pointrange(data=left_join(filter(comp.emm, trait=="prop.hybrid"), hybrid.test),
                  aes(y=estimate, ymax=conf.high, ymin=conf.low, shape=p.adj < 0.05)) + 
  scale_shape_manual("Different from 50%", values=c(1, 19), labels=c("no","yes"))+
  geom_text(data=hybrid.test, aes(y=1.1, label=crosses))+
  #geom_text(data=hybrid.test, aes(y=1.1, label=alive))+
  scale_fill_brewer(palette="Paired", guide="none") + 
  scale_y_continuous(breaks=(0:4)/4, labels=~scales::percent(.,accuracy=1))+ 
  labs(x="Maternal species", y="Hybrids from competitive pollinations") + sp_theme
```

Tested whether the proportion of hybrids significantly differed from 50%, accounting for multiple comparisons. Violin plots show proportions for each plant x plant cross. The number of plant x plant crosses indicated above the plot.

# Phylogenetic distance

Load *Schiedea* tree inferred with RAxML from Willyard et al. nuclear dataset (no chloroplast data, 1 accession per species).

```{r phylodist}
library(ape)
library(RColorBrewer)
#devtools::install_github('gastonstat/arcdiagram')
library(igraph)
library(arcdiagram)

sp_pal <- set_names(brewer.pal(length(sp_pairs), "Paired"), sp_pairs)
tree <- read.nexus("data/RAxML_1ps_nocp_bipartitionsBranchLabels.sch.short.nobs.nex")
plot(tree, tip.color = sp_pal[tree$tip.label])

par(mfrow=c(1,2))
par(mar=c(1.5,0,1.5,0))
tree_pairs <- tree %>% keep.tip(sp_pairs)
plot(tree_pairs, tip.color = sp_pal[tree_pairs$tip.label])
par(mar=c(0,0,0,0))
sp_pairs %>% matrix(ncol=2, byrow=T) %>% 
  arcplot(show.labels=F, horizontal=F, ordering=tree_pairs$tip.label, 
          col.arcs=sp_pal[(1:5)*2], col.nodes = sp_pal)
par(mfrow=c(1,1))

hybrid.phylodist <- comp.emm %>% left_join(select(hybrid.test, momsp, dadsp, -type)) %>% 
  mutate(pd = map2_dbl(as.character(momsp), as.character(dadsp), ~cophenetic(tree_pairs)[.x,.y])) %>% 
  select(-df)

hybrid.phylodist.test <- hybrid.phylodist %>% group_by(trait, type) %>% nest() %>% 
  mutate(model = map(data, ~lm(estimate ~ pd, data=.x)),
         anova = map(model, ~tidy(car::Anova(.x)))) %>% 
  select(-data, -model) %>% unnest(anova) %>% filter(term !="Residuals") 

ggplot(left_join(hybrid.phylodist, hybrid.phylodist.test), aes(x=pd, y=estimate)) + 
  facet_grid(trait~type, labeller=as_labeller(c(propnames, typenames)))+
  geom_smooth(aes(linetype=paste(p.value < 0.05, type=="intra")), method="lm", se=F, color="grey") +
  geom_pointrange(aes(ymax=conf.high, ymin=conf.low, color=momsp)) + scale_color_brewer(palette="Paired") + 
  scale_y_continuous(limits = c(0,1), breaks=(0:4)/4, labels=~scales::percent(.,accuracy=1), expand=c(0,0))+ 
  scale_linetype_manual(values=c(2,0,1,0), guide="none") + scale_alpha_manual(breaks=c(T,F), values=c(0.5, 1))+
  labs(x="Phylogenetic distance (mutations/site)", color="Maternal species") + theme_bw() +
  theme(axis.title.y=element_blank(), axis.text=element_text(color="black"), legend.position="top",
        panel.grid.minor = element_blank(), panel.grid.major.x=element_blank(), panel.spacing = unit(1, "lines"))
```

Regressions against the phylogenetic distance of each species pair. The solid regression line is statistically significant (P = 0.001). Note that for reference, data for intraspecific crosses is plotted with the phylogenetic distance to its paired species, which doesn't make great sense.

# Style length

The range of phylogenetic distances and style lengths in the study. The species in each pair with a longer style is given a lighter color.

```{r styles}
hybrid.phylodist %>% filter(trait=="prop.hybrid",  type=="mixed") %>% left_join(styles) %>% 
  ggplot(aes(x=pd, y=style_length_mm, color=momsp, shape=mating)) + 
  geom_pointrange(aes(ymin=style_length_mm-se, ymax=style_length_mm+se)) + 
  scale_color_brewer(palette="Paired") + coord_cartesian(xlim=c(0, NA), ylim=c(0,NA)) +
  labs(x="Phylogenetic distance (mutations/site)", y="Style length (mm)", color="Species", shape="Mating") + 
  theme(legend.position = "top")

ggplot(left_join(comp.emm, styles), aes(x=style_length_mm, y=estimate)) + 
  facet_grid(trait~type, labeller=as_labeller(c(propnames, typenames)))+
  geom_line(aes(group=pair, color=momsp))+
  geom_pointrange(aes(ymax=conf.high, ymin=conf.low, color=momsp, shape=mating)) + scale_color_brewer(palette="Paired") + 
  scale_y_continuous(limits = c(0,1), breaks=(0:4)/4, labels=~scales::percent(.,accuracy=1), expand=c(0,0))+ 
  labs(x="Style length (mm)", color="Maternal species", shape="Mating") + theme_bw() +
  theme(axis.title.y=element_blank(), axis.text=element_text(color="black"), legend.position="top",
        panel.grid.minor = element_blank(), panel.grid.major.x=element_blank(), panel.spacing = unit(1, "lines"))
```

Lines show the effect of style length within each pair. Style length was not a significant predictor of any variable for any cross category in models that also included phylogenetic distance.

