---
title: "Schiedea hookeri/kaalae RIMs"
author: "[John Powers](http://sites.uci.edu/powers/)"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

```{r setup, message = FALSE}
library(tidyverse)
library(RColorBrewer)

library(glmmTMB)
library(broom)
library(emmeans)

library(knitr)
knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, fig.path = "figures/", dev = "svglite", fig.width=5, fig.height=4, width=999)
options(digits=4, knitr.kable.NA = "") # for kables

load("data/rims_data.rda")

sxc.labels <- set_names(c("K intrapop", "K interpop", "K x H", "H x K", "H interpop", "H intrapop"), sxc.levels)
F1.col <- set_names(c(brewer.pal(3,"Greens")[2:3],brewer.pal(9,"Oranges")[5], brewer.pal(9,"Set1")[7], brewer.pal(3,"Purples")[3:2]), sxc.levels)
```

# Datasets

For seed production (capsule formation and viable seeds), the generation refers to the zygote generation (e.g. F2 seed grown from an F1 plant), otherwise the plant that produced the trait (e.g. F2 pollen is from F2 plants). This convention isn't always carried over to the name of the dataset for historical reasons.

```{r datasets}
traits %>% filter(generation !="P") %>% select(c(generation,ylab, yvar, data, type)) %>% kable
```

# F1 Generation

## Sample sizes

These sample sizes are the number of rows in the dataset. The rows mean different things in each dataset:

* seeds = seeds from one pollination
* germination = seeds from one pollination (multiple pots are pooled)
* vegbiomass = one offspring plant, if it survived
* survflr = one offspring plant
* pollen = one pollen collection
* inflobiomass = one offspring plant, if it flowered

```{r sampsize}
F1 <- set_names(datanames[1:6]); F1[[6]] <- "inflobiomass.sum"

map_dfr(F1, ~ tally(get(.)), .id="dataset") %>% 
  pivot_wider(names_from=dataset, values_from=n) %>% kable(caption="F1 datasets totals")

map_dfr(F1, ~ tally(group_by(get(.), crosstype, momsp, dadsp)), .id="dataset") %>% 
  pivot_wider(names_from=dataset, values_from=n) %>% kable(caption="F1 datasets by crosstype and species")

map_dfr(F1, ~ tally(group_by(get(.), crosstype, momsp, dadsp, mompop, dadpop)), .id="dataset") %>% 
  pivot_wider(names_from=dataset, values_from=n) %>% kable(caption="F1 datasets by crosstype and population")
```

## Models

Expand this code to see the model specifications. The models have the following settings:

* response variable: for details see the [metadata](https://docs.google.com/spreadsheets/d/1f487VMTsYzYGn0_daogtTF0nNE7ntoqd0Yh4h9tgyDE/edit#gid=2002881351)
* sxc: the combined maternal plant species and crosstype (within population, between population, hybrid)
* collect: the number of days from planting to harvest, to account for growth during harvest period
* mompid, dadpid: random effects of the maternal and paternal plant genotypes (clones are treated as the same genotype)
* data: for details and sample sizes see tables above
* family: gaussian (the default, for continuous variables), truncated_nbinom1 (quasi-Poisson truncated at zero for nonzero seed counts), betabinomial (for binary variables)


```{r models}
F1.mod <- list(
  firstflower =          glmmTMB(firstflower ~          sxc +           (1|mompid) + (1|dadpid), data=survflr),
  firstinflo.biomass.g = glmmTMB(firstinflo.biomass.g ~ sxc + collect + (1|mompid) + (1|dadpid), data=survflr),
  inflo.biomass.g =      glmmTMB(inflo.biomass.g  ~     sxc + collect + (1|mompid) + (1|dadpid), data=inflobiomass.sum),
  flower.number =        glmmTMB(flower.number  ~       sxc + collect + (1|mompid) + (1|dadpid), data=inflobiomass.sum),
  veg.biomass.g =        glmmTMB(veg.biomass.g ~        sxc + collect + (1|mompid) + (1|dadpid), data=vegbiomass),
  total.per.anther =     glmmTMB(total.per.anther ~     sxc +           (1|mompid) + (1|dadpid), data=pollen),
  viable.per.anther =    glmmTMB(viable.per.anther ~    sxc +           (1|mompid) + (1|dadpid), data=pollen),
  viable.seeds =         glmmTMB(viable.seeds ~         sxc +           (1|mompid) + (1|dadpid), data=filter(seeds, viable.seeds>0),
                                 family="truncated_nbinom1"),
  prop.germ =            glmmTMB(prop.germ      ~ sxc + (1|mompid) + (1|dadpid), family="betabinomial", data=germination, weights = planted),
  alive =                glmmTMB(alive          ~ sxc + (1|mompid) + (1|dadpid), family="betabinomial", data=survflr),
  flowered =             glmmTMB(flowered       ~ sxc + (1|mompid) + (1|dadpid), family="betabinomial", data=filter(survflr, alive)),
  prop.viable =          glmmTMB(prop.viable    ~ sxc + (1|mompid) + (1|dadpid), family="betabinomial", data=pollen, weights = total),
  capsule.formed =       glmmTMB(capsule.formed ~ sxc + (1|mompid) + (1|dadpid), family="betabinomial", data=seeds))
save(F1.mod, file="data/F1mod.rda")
#load("data/F1mod.rda")
```

## Inference 

P-values for the overall effect of species and crosstype, or collection date, on the trait.

```{r inference}
F1.test <- map_dfr(F1.mod, ~car::Anova(.) %>% tidy(), .id="trait") 
F1.test %>% left_join(traits %>% filter(generation=="F1") %>% select(trait=yvar, step)) %>% 
  arrange(step) %>% select(-step) %>% 
  mutate(term=recode(term, sxc = "species and crosstype", collect= "collection date"),
         trait = recode(trait, !!!set_names(traits$ylab[-1], traits$yvar[-1])) %>% 
               str_remove(fixed(" (g)")),
         p.value=format(p.value,digits=2)) %>% 
  kable(digits=1, caption="ANOVA of GLMM (Type III Wald chisquare tests)")
  
F1.emm <-  map_dfr(F1.mod, ~emmeans(., ~sxc) %>% summary(type="response") %>%  #average over collect date
                             tidy %>% rename(estimate=any_of(c("response","prob"))), .id="trait") 
```

## Plots

Estimated marginal means with 95% CI are shown as dots. For the continuous variables, these show violin plots of the raw data with median and first and third quartiles. For binary variables, colored dots show the proportions of each population x population combination.

```{r f1plot}
plot_F1 <- function(emm, data, yvar, ylab, type="continuous") {
  data <- get(data)
  if(type=="continuous") {
    p <- ggplot(data, aes_string(y=yvar, x="sxc", fill="sxc")) +
      geom_violin(draw_quantiles = (1:3)/4) + 
      geom_pointrange(data=filter(emm, trait==yvar), aes(y=estimate, ymax=conf.high, ymin=conf.low)) + 
      scale_fill_manual(values=F1.col) + guides(fill="none") +
      scale_y_continuous(name = ylab, expand = expansion(mult=c(0, 0.05)), limits=c(0,NA))
  } else if(type=="binary") {
    p <- data %>% drop_na(all_of(yvar)) %>% 
      group_by(sxc, mompop, dadpop) %>% summarize(across(all_of(yvar), ~sum(.)/n()), n=n(), .groups="drop") %>% 
      ggplot(aes_string(y=yvar, x="sxc", color="sxc")) +
      geom_point(size=4)+
      geom_pointrange(data=filter(emm, trait==yvar) %>% 
                        mutate(conf.low  = ifelse(conf.high-conf.low>0.9,NA,conf.low), 
                               conf.high = ifelse(conf.high-conf.low>0.9,NA,conf.high)), 
                      aes(y=estimate, ymax=conf.high, ymin=conf.low), color="black") + 
      scale_color_manual(values=F1.col) + guides(color="none") +
      scale_y_continuous(ylab, limits=c(0,1), labels=~scales::percent(.,accuracy=1))
  }
  p <- p + scale_x_discrete("Cross", labels = sxc.labels) + 
    theme_minimal() + theme(panel.grid.major.x = element_blank())
  print(p)
}

traits %>% filter(generation=="F1") %>% select(data, yvar, ylab, type) %>% 
  pwalk(plot_F1, emm=F1.emm)
```

## Flower biomass regression

```{r infloreg}
summary(biomass.flrs.lm)

ggplot(inflobiomass, aes(x=inflo.biomass.g, y=flrs, color=cross)) + 
  geom_point() +  geom_smooth(method="lm", formula = y ~ poly(x,2), se=F)+
  geom_rug(data=inflobiomass.sum, aes(x=inflo.biomass.g/inflo, y=10), sides="b") + 
  scale_x_log10() + scale_y_log10() + 
  scale_color_manual(values=set_names(F1.col, c("KKwithin","KK", "KH", "HK", "HH", "HHwithin"))[-c(1,6)]) +
  theme_minimal() + labs(y="Flower number", x="Inflorescence biomass (g)", color="Cross")

ggplot(inflobiomass, aes(x=inflo.biomass.g, y=flrs/inflo.biomass.g, color=cross)) + 
  geom_point() +  geom_smooth(se=F, span=0.8) +
  scale_x_log10() + scale_y_log10() + 
  scale_color_manual(values=set_names(F1.col, c("KKwithin","KK", "KH", "HK", "HH", "HHwithin"))[-c(1,6)]) +
  theme_minimal() + labs(y="Flowers per g biomass", x="Inflorescence biomass (g)", color="Cross")

ggplot(inflobiomass.sum, aes(y=inflo, x=sxc, fill=sxc)) +
  geom_boxplot() + geom_violin(fill=NA) + 
  scale_fill_manual(values=F1.col) + guides(fill="none") +
  scale_y_continuous("Inflorescences", expand = expansion(mult=c(0, 0.05)), limits=c(0,NA))+
  scale_x_discrete("Cross", labels = sxc.labels) + 
  theme_minimal() + theme(panel.grid.major.x = element_blank())

ggplot(inflobiomass.sum, aes(y=inflo.biomass.g/inflo, x=sxc, fill=sxc)) +
  geom_boxplot() + geom_violin(fill=NA) +  
  scale_fill_manual(values=F1.col) + guides(fill="none") +
  scale_y_continuous("Biomass per inflorescence (g)", expand = expansion(mult=c(0, 0.05)), limits=c(0,NA))+
  scale_x_discrete("Cross", labels = sxc.labels) + 
  theme_minimal() + theme(panel.grid.major.x = element_blank())

ggplot(inflobiomass.sum, aes(y=flower.number/inflo, x=sxc, fill=sxc)) +
  geom_boxplot() + geom_violin(fill=NA) + 
  scale_fill_manual(values=F1.col) + guides(fill="none") +
  scale_y_continuous("Flowers per inflorescence", expand = expansion(mult=c(0, 0.05)), limits=c(0,NA))+
  scale_x_discrete("Cross", labels = sxc.labels) + 
  theme_minimal() + theme(panel.grid.major.x = element_blank())

```

## Fitness and RI

```{r fitness, fig.width=6}
F1.emm %>%  left_join(traits %>% filter(generation=="F1") %>% select(trait=yvar, step)) %>% arrange(step) %>% 
  mutate(est.se = paste(round(estimate,2.5-log10(estimate)), "\U00B1", round(std.error,2.5-log10(estimate))),
                  trait = recode(trait, !!!set_names(traits$ylab[-1], traits$yvar[-1])),
                  sxc = recode(sxc, !!!sxc.labels)) %>% 
  select(trait, sxc, est.se) %>% pivot_wider(names_from = sxc, values_from=est.se) %>% 
  kable(caption="Absolute fitness components with estimated marginal mean and standard error")

#TODO to get confidence intervals of RI ratio, may need mratios or alt method here:
# https://stats.stackexchange.com/questions/16349/how-to-compute-the-confidence-interval-of-the-ratio-of-two-normal-means
calc_RI <- function(h, b) 1 - 2 *  h/(h + b) #method of Sobel and Chen 2014 for hybrids vs. between-population intraspecific fitness
calc_ID <- function(w, b) 1 - w/b
highlow <- function(fn, a, b) { #method of Sobel and Streisfeld 2015 Fig. S1
  combos <- c(fn(a[1], b[1]), fn(a[1],b[2]), fn(a[2],b[1]), fn(a[2], b[2]))
  set_names(c(min(combos), max(combos)), c("low","high"))
}

F1.RI <- F1.emm %>% separate(sxc, into=c("momsp","crosstype")) %>% 
  filter(crosstype !="within", trait != "firstflower") %>% #TODO firstflower is backwards fitness
  select(trait, momsp, crosstype, estimate, conf.low, conf.high) %>% 
  pivot_wider(names_from=crosstype, values_from=c(estimate, conf.low, conf.high)) %>% 
  rename_with(~ str_remove(.,"estimate_|conf.")) %>% rowwise() %>% 
  mutate(across(c(between,hybrid), ~ ./max(between,hybrid)),
         RI = calc_RI(h=hybrid, b=between)) %>%
  mutate(RI.range = list(highlow(fn=calc_RI, a=c(low_hybrid, high_hybrid), b=c(low_between, high_between))), 
         .keep="unused") %>% unnest_wider(RI.range) %>% 
  left_join(traits %>% filter(generation=="F1") %>% select(trait=yvar, step, sex)) %>% arrange(step) %>% 
  mutate(trait = recode(trait, !!!set_names(traits$ylab[-1], traits$yvar[-1])) %>% 
           str_remove(fixed(" (g)")) %>% factor() %>% fct_reorder(step)) 
  
abs_contribution <- function(RI) { #method of Ramsey et al. 2003 for absolute contribution to RI
  AC <- RI
  for(n in 1:length(RI)) {
    AC[[n]] <- RI[[n]] * (1 - sum(AC[1:n-1]))
  }
  return(AC)
}
#cumulative reproductive isolation 
F1.CRI.m <- F1.RI %>% filter(str_detect(sex,"m")) %>% select(trait, momsp, RI) %>% 
              pivot_wider(names_prefix="RI_", names_from=momsp, values_from=RI) %>% 
              mutate(across(starts_with("RI"), abs_contribution, .names="AC{.col}"), 
                     across(starts_with("AC"), ~ . /sum(.),  .names="R{.col}"),
                     across(starts_with("AC"), cumsum, .names="T{.col}"),
                     .keep = "unused")

F1.RI %>% select(trait, momsp, between, hybrid, RI) %>% 
  pivot_wider(names_from=momsp, values_from=c(between, hybrid, RI)) %>% 
  left_join(F1.CRI.m) %>% 
  select(trait, KK = between_kaal, KH = hybrid_kaal, K_RI = RI_kaal, K_AC = ACRI_kaal, K_RC = RACRI_kaal, 
                HH = between_hook, HK = hybrid_hook, H_RI = RI_hook, H_AC = ACRI_hook, H_RC = RACRI_hook) %>% 
  kable(caption="Relative fitness components of hybrids vs. maternal parent. RI, AC, RC follow Sobel & Chen 2014, Ramsey et al. 2003", digits=3)

ggplot(F1.RI, aes(y=fct_rev(trait), x=RI, color=momsp)) + 
  geom_vline(xintercept=0, linetype=2) + geom_vline(xintercept=-1)+ geom_vline(xintercept=1)+ 
  geom_linerange(aes(xmin=low, xmax=high), position=position_dodge(width=0.5)) + 
  geom_point(shape=17, size=2, position=position_dodge(width=0.5)) +
  scale_color_manual(values = set_names(F1.col[c(2,5)], c("kaal","hook")), labels=c("S. kaalae","S. hookeri")) +
  scale_x_continuous(limits=c(-1.1,1.1), breaks=c(-1,-0.5,0,0.5,1),
                     labels=c("-1\nheterosis","-0.5","0\nno\nbarrier","0.5","1\ncomplete\nisolation")) + 
  theme_minimal() + theme(legend.position = "top", legend.text = element_text(face="italic"))+
  labs(x="Reproductive isolation", y="Fitness component", color="Maternal parent")

F1.CRI.m %>% summarize(across(starts_with("AC"), sum)) %>% rename_with(~str_remove(.,"ACRI_")) %>% 
  kable(caption="Total male reproductive isolation = sum of absolute contributions (AC) to RI", digits=3)

F1.CRI.m %>% pivot_longer(starts_with("TAC"), names_prefix="TACRI_", names_to="momsp", values_to="TAC") %>% 
  ggplot(aes(y=fct_rev(trait), x=TAC, color=momsp)) + 
  geom_vline(xintercept=0, linetype=2) + geom_vline(xintercept=1)+ 
  geom_path(aes(group=momsp))+
  geom_point(shape=17, size=2) +
  scale_color_manual(values = set_names(F1.col[c(2,5)], c("kaal","hook")), labels=c("S. kaalae","S. hookeri")) +
  scale_x_continuous(limits=c(NA, 1.1), breaks=c(0,0.5,1),
                     labels=c("0\nno\nbarrier","0.5","1\ncomplete\nisolation")) +
  theme_minimal() + theme(legend.position = "top", legend.text = element_text(face="italic"))+
  labs(x="Cumulative reproductive isolation", y="Fitness component", color="Maternal parent")
```

### Distribution of cumulative fitness

```{r totalRI_simulation, fig.height=6, fig.width=8}
#dataname = name of the dataset, yvar = the fitness trait, N = number of trials for each species and crosstype
sample_trait <- function(dataname, yvar, N) {
  get(dataname) %>% select(sxc, y=!!yvar) %>% 
    drop_na(y) %>% #drop missing trait values (various reasons for these)
    group_by(sxc) %>% nest() %>% #subset data by the species and crosstype (between/within/hybrid)
    mutate(sampled = map(data, slice_sample, n=N, replace=T)) %>% #sample each subset N times with replacement
    select(-data) %>% unnest(sampled) %>% arrange(sxc) %>% mutate(trial=row_number()) #give each trial an index
}

set.seed(1)
ntrials <- 10000
F1.cumul.fitness <- traits %>% 
  filter(generation=="F1", str_detect(sex,"m")) %>% select(data, yvar) %>% # independent multiplicative traits
  mutate(samp = pmap(., sample_trait, N = ntrials)) %>% # sample each trait N times per species and crosstype
  select(-data) %>% unnest(samp) %>% # expand sampled dataset to full size
  pivot_wider(id_cols=c(sxc, trial), names_from=yvar, values_from=y) %>% # line up the trials for each variable
  mutate(fitness_capsule.formed   = capsule.formed,
         fitness_viable.seeds     = capsule.formed * viable.seeds,
         fitness_prop.germ        = capsule.formed * viable.seeds * prop.germ,
         fitness_alive            = capsule.formed * viable.seeds * prop.germ * alive,
         fitness_flowered         = capsule.formed * viable.seeds * prop.germ * alive * flowered,
         fitness_flower.number    = capsule.formed * viable.seeds * prop.germ * alive * flowered * flower.number,
         fitness_total.per.anther = capsule.formed * viable.seeds * prop.germ * alive * flowered * flower.number * total.per.anther,
         fitness_prop.viable      = capsule.formed * viable.seeds * prop.germ * alive * flowered * flower.number * total.per.anther * prop.viable) 

F1.cumul.fitness %>% select(sxc, starts_with("fitness")) %>% 
  pivot_longer(starts_with("fitness"), names_to="stage", names_prefix = "fitness_") %>% 
  ggplot(aes(x=value, color=sxc)) + facet_wrap(vars(stage), scales="free_x", labeller=as_labeller(deframe(select(traits, yvar, ylab))))+
  geom_freqpoly(bins=30, boundary=1, position=position_identity(), size=1) +
  scale_color_manual("Crosstype", values=F1.col, labels=sxc.labels) + 
  scale_y_continuous(labels = ~scales::percent(./ntrials)) + scale_x_continuous(limits=c(1,NA)) + theme_minimal() + 
  labs(y=paste0("Frequency of outcome (",ntrials," trials)"), 
       x="F1 flowers or F1 viable pollen grains resulting from one P x P pollination",
       title = "Distribution of female and male cumulative F1 fitness (if nonzero)") 

F1.cumul.fitness %>% group_by(sxc) %>% summarize(prop.zero=mean(fitness_flower.number>0)) %>% 
ggplot(aes(y=prop.zero, x=sxc, color=sxc)) + 
  geom_point(size=2)+ 
  scale_color_manual(values=F1.col, guide="none") + 
  scale_y_continuous(limits=c(0,1), labels=scales::percent) + scale_x_discrete(labels=sxc.labels) + theme_minimal() + 
  labs(y=paste0("Nonzero F1 flowers resulting from one P x P pollination (",ntrials," trials)"), x="Crosstype",
       title = "Frequency of female cumulative F1 fitness greater than zero") 
```

```{r totalRI_means, fig.height=6, fig.width=6.5}
trials_to_average <- 50
F1.cumul.isolation <- F1.cumul.fitness %>% 
  group_by(sxc, meanblock = ceiling(trial / trials_to_average)) %>% #group into blocks of trials_to_average trials
  summarize(across(starts_with("fitness"), mean)) %>% #take the mean fitness for each block and crosstype
  pivot_longer(starts_with("fitness"), names_prefix = "fitness_", names_to="stage", values_to="fitness") %>% 
  separate(sxc, into=c("momsp","crosstype")) %>% 
  pivot_wider(names_from=crosstype, values_from=fitness) %>% 
  mutate(stage = factor(stage, levels=unique(traits$yvar)),
         RI = calc_RI(h=hybrid, b=between),
         ID = calc_ID(w=within, b=between)) 

F1.CRI.TACmf <- F1.CRI.m %>% 
  pivot_longer(starts_with("TAC"), names_prefix="TACRI_", names_to="momsp", values_to="TAC") %>% 
  mutate(stage = recode(trait, !!!set_names(traits$yvar[-1], traits$ylab[-1]))) %>% 
  filter(stage %in% unique(F1.cumul.isolation$stage)) %>% 
  rename(RI=TAC)

ggplot(F1.cumul.isolation, aes(y=fct_rev(stage), color=momsp, x=RI)) + 
  geom_vline(xintercept=0, linetype=2) + geom_vline(xintercept=-1)+ geom_vline(xintercept=1)+ 
  geom_violin(draw_quantiles=(1:3)/4, position=position_dodge(width=0.5)) + 
  geom_path(aes(group=momsp), data=F1.CRI.TACmf, position=position_dodge(width=0.5)) +
  geom_point(data=F1.CRI.TACmf, position=position_dodge(width=0.5), size=2, shape=17) +
  scale_y_discrete(labels=deframe(select(traits, yvar, ylab))) +
  scale_color_manual(values = set_names(F1.col[c(2,5)], c("kaal","hook")), labels=c("S. kaalae","S. hookeri")) +
  scale_x_continuous(limits=c(-1.1,1.1), breaks=c(-1,-0.5,0,0.5,1),
                     labels=c("-1\nheterosis","-0.5","0\nno\nbarrier","0.5","1\ncomplete\nisolation")) + 
  theme_minimal() + theme(legend.position = "top", legend.text = element_text(face="italic")) +
  labs(x="Cumulative reproductive isolation", y="Fitness component", color="Maternal parent")
```


## Inbreeding depression

```{r inbred}
F1.inbred <- F1.emm %>% separate(sxc, into=c("momsp","crosstype")) %>% 
  filter(crosstype !="hybrid", trait != "firstflower") %>% #TODO firstflower is backwards fitness
  select(trait, momsp, crosstype, estimate, conf.low, conf.high) %>% 
  pivot_wider(names_from=crosstype, values_from=c(estimate, conf.low, conf.high)) %>% 
  rename_with(~ str_remove(.,"estimate_|conf.")) %>% rowwise() %>% 
  mutate(across(c(between,within), ~ ./max(between,within)),
         ID = calc_ID(w=within, b=between)) %>%
  mutate(ID.range = list(highlow(fn=calc_ID, a=c(low_within, high_within), b=c(low_between, high_between))), 
         .keep="unused") %>% unnest_wider(ID.range) %>% 
  left_join(traits %>% filter(generation=="F1") %>% select(trait=yvar, step, sex)) %>% arrange(step) %>% 
  mutate(trait = recode(trait, !!!set_names(traits$ylab[-1], traits$yvar[-1])) %>% 
           str_remove(fixed(" (g)")) %>% factor() %>% fct_reorder(step)) 

F1.inbred %>% select(trait, momsp, between, within, ID) %>% 
  pivot_wider(names_from=momsp, values_from=c(between, within, ID)) %>% 
  select(trait, between_kaal, within_kaal, ID_kaal, between_hook, within_hook, ID_hook) %>% 
  kable(caption="Inbreeding depression \u03B4 = 1 - (within pop / between pop)", digits=3)

ggplot(F1.inbred, aes(y=fct_rev(trait), x=ID, color=momsp)) + 
  geom_vline(xintercept=0, linetype=2) +  
  geom_linerange(aes(xmin=low, xmax=high), position=position_dodge(width=0.5)) + 
  geom_point(shape=17, size=2, position=position_dodge(width=0.5)) +
  scale_color_manual(values = set_names(F1.col[c(2,5)], c("kaal","hook")), labels=c("S. kaalae","S. hookeri")) +
  scale_x_continuous(limits=c(-1.1,1.1), breaks=c(-1,-0.5,0,0.5,1),
                     labels=c("-1\noutcrossing\ndepression","-0.5","0\nneutral","0.5","1\ninbreeding\ndepression")) +
  theme_minimal() + theme(legend.position = "top", legend.text = element_text(face="italic"))+
  labs(x="Inbreeding depression \u03B4", y="Fitness component", color="Species")

# Cumulative inbreeding depression
F1.CID.m <- F1.inbred %>% filter(str_detect(sex,"m")) %>%  select(trait, momsp, ID) %>% 
              pivot_wider(names_prefix="ID_", names_from=momsp, values_from=ID) %>% 
              mutate(across(starts_with("ID"), abs_contribution, .names="AC{.col}"), 
                     across(starts_with("AC"), ~ . /sum(.),  .names="R{.col}"),
                     across(starts_with("AC"), cumsum, .names="T{.col}"),
                     .keep = "unused")

F1.inbred %>% select(trait, momsp, between, within, ID) %>% 
  pivot_wider(names_from=momsp, values_from=c(between, within, ID)) %>% 
  left_join(F1.CID.m) %>% 
  select(trait, Kb = between_kaal, Kw = within_kaal, K_ID = ID_kaal, K_AC = ACID_kaal, K_RC = RACID_kaal, 
                Hb = between_hook, Hw = within_hook, H_ID = ID_hook, H_AC = ACID_hook, H_RC = RACID_hook) %>% 
  kable(caption="Relative fitness components of within vs. between population crosses", digits=3)

F1.CID.m %>% summarize(across(starts_with("AC"), sum)) %>% rename_with(~str_remove(.,"ACID_")) %>% 
  kable(caption="Total male inbreeding depression = sum of absolute contributions (AC) to ID", digits=3)

F1.CID.m %>% pivot_longer(starts_with("TAC"), names_prefix="TACID_", names_to="momsp", values_to="TAC") %>% 
  ggplot(aes(y=fct_rev(trait), x=TAC, color=momsp)) + 
  geom_vline(xintercept=0, linetype=2) + geom_vline(xintercept=1)+ 
  geom_path(aes(group=momsp))+
  geom_point(shape=17, size=2) +
  scale_color_manual(values = set_names(F1.col[c(2,5)], c("kaal","hook")), labels=c("S. kaalae","S. hookeri")) +
  scale_x_continuous(limits=c(NA, 1.1), breaks=c(0,0.5,1),
                     labels=c("0\nneutral","0.5","1\ninbreeding\ndepression")) +
  theme_minimal() + theme(legend.position = "top", legend.text = element_text(face="italic"))+
  labs(x="Cumulative inbreeding depression", y="Fitness component", color="Species")
```

```{r totalID_means, fig.height=6, fig.width=6.5}
# Distribution of ID
F1.CID.TACmf <- F1.CID.m %>% 
  pivot_longer(starts_with("TAC"), names_prefix="TACID_", names_to="momsp", values_to="TAC") %>% 
  mutate(stage = recode(trait, !!!set_names(traits$yvar[-1], traits$ylab[-1]))) %>% 
  filter(stage %in% unique(F1.cumul.isolation$stage)) %>% 
  rename(ID=TAC)

ggplot(F1.cumul.isolation, aes(y=fct_rev(stage), color=momsp, x=ID)) + 
    geom_vline(xintercept=0, linetype=2) + geom_vline(xintercept=-1)+ geom_vline(xintercept=1)+ 
    geom_violin(draw_quantiles=(1:3)/4, position=position_dodge(width=0.9)) +
    geom_path(aes(group=momsp), data=F1.CID.TACmf, position=position_dodge(width=0.9)) +
    geom_point(data=F1.CID.TACmf, position=position_dodge(width=0.9), size=2, shape=17) +
    scale_y_discrete(labels=deframe(select(traits, yvar, ylab))) +
    scale_color_manual(values = set_names(F1.col[c(2,5)], c("kaal","hook")), labels=c("S. kaalae","S. hookeri")) +
    scale_x_continuous(limits=c(-1.1,1.1), breaks=c(-1,-0.5,0,0.5,1),
                     labels=c("-1\noutcrossing\ndepression","-0.5","0\nneutral","0.5","1\ninbreeding\ndepression")) + 
    theme_minimal() + theme(legend.position = "top", legend.text = element_text(face="italic")) +
    labs(x="Cumulative inbreeding depression", y="Fitness component", color="Species")
```


# F2 Generation

## Sample sizes

These sample sizes are the number of rows in the dataset. The rows mean different things in each dataset:

* f1seeds = seeds from one pollination
* f1seeds = the mass of ten seeds
* f2pollen = one pollen collection
* f2seeds = seeds from one pollination

```{r f2sampsize}
F2 <- set_names(datanames[7:10])

map_dfr(F2, ~ tally(get(.)), .id="dataset") %>% 
  pivot_wider(names_from=dataset, values_from=n) %>% kable(caption="F2 datasets totals")

map_dfr(F2, ~ tally(group_by(get(.), generation, momdadcross)), .id="dataset") %>% 
  pivot_wider(names_from=dataset, values_from=n) %>% kable(caption="F2 datasets by cross")
```

## Models

```{r f2models}
F2.mod <- list(
  seed.mass.mg =         glmmTMB(seed.mass.mg ~      momdadcross, data=f1seedmass),
  total.per.anther =     glmmTMB(total.per.anther ~  momdadcross, data=f2pollen),
  viable.per.anther =    glmmTMB(viable.per.anther ~ momdadcross, data=f2pollen),
  viable.seeds =         glmmTMB(viable.seeds ~      momdadcross, data=filter(f1seeds, viable.seeds>0), family="truncated_nbinom1"),
  prop.viable =          glmmTMB(prop.viable ~       momdadcross, family="betabinomial", data=f2pollen, weights = total),
  capsule.formed =       glmmTMB(capsule.formed ~    momdadcross, family="betabinomial", data=f1seeds))
```


## Inference

```{r f2inference}
F2.test <- map_dfr(F2.mod, ~car::Anova(.) %>% tidy(), .id="trait") 
F2.test %>% left_join(traits %>% filter(generation=="F2") %>% select(trait=yvar, step)) %>% 
  arrange(step) %>% select(-step) %>% 
  mutate(trait = recode(trait, !!!set_names(traits$ylab[-1], traits$yvar[-1])),
         p.value=format(p.value,digits=2)) %>% 
  kable(digits=1, caption="ANOVA of GLMM (Type III Wald chisquare tests)")
  
F2.emm <-  map_dfr(F2.mod, ~emmeans(., ~momdadcross) %>% summary(type="response") %>%  
                             tidy %>% rename(estimate=any_of(c("response","prob"))), .id="trait") 
```


## Plots

Estimated marginal means with 95% CI are shown as dots. For the continuous variables, these show violin plots of the raw data with median and first and third quartiles.

### All crosses

```{r f2plot}
momdadcross.levels <- c("H x H","H x HK","H x KH","H x K","HK x H","HK x HK","HK x KH","HK x K","KH x H","KH x HK","KH x KH","KH x K","K x H","K x HK","K x KH","K x K")
momdadcross.labels <- momdadcross.levels %>% str_replace_all("\\s","\n") %>% set_names(momdadcross.levels)

plot_F2 <- function(emm, data, yvar, ylab, type="continuous") {
  data <- get(data) %>% mutate(momdadcross = factor(as.character(momdadcross), levels=momdadcross.levels))
  if(type=="continuous") {
    p <- ggplot(data, aes_string(y=yvar, x="momdadcross", fill="momcross")) + 
      geom_violin(draw_quantiles = (1:3)/4) + 
      geom_pointrange(data=filter(emm, trait==yvar) %>% 
                        separate(momdadcross, into=c("momcross","dadcross"), sep=" x ", remove=F) %>% 
                        mutate(momdadcross = factor(as.character(momdadcross), levels=momdadcross.levels)),
                      aes(y=estimate, ymax=conf.high, ymin=conf.low),color="black") + 
      scale_fill_manual(values=set_names(F1.col[2:5],c("K","KH","HK","H")))+
      guides(fill="none") +
      scale_y_continuous(name = ylab, expand = expansion(mult=c(0, 0.05)), limits=c(0,NA))
  } else if(type=="binary") {
    p <- ggplot(data, aes_string(y=yvar, x="momdadcross", color="momcross")) + 
      geom_pointrange(data=filter(emm, trait==yvar) %>% 
                        separate(momdadcross, into=c("momcross","dadcross"), sep=" x ", remove=F) %>% 
                        mutate(momdadcross = factor(as.character(momdadcross), levels=momdadcross.levels)) %>% 
                        mutate(conf.low  = ifelse(conf.high-conf.low>0.9,NA,conf.low), 
                               conf.high = ifelse(conf.high-conf.low>0.9,NA,conf.high)), 
                      aes(y=estimate, ymax=conf.high, ymin=conf.low)) + 
      scale_color_manual(values=set_names(F1.col[2:5],c("K","KH","HK","H")))+
      guides(color="none") +
      scale_y_continuous(ylab, limits=c(0,1), labels=~scales::percent(.,accuracy=1))
  }
  p <- p + scale_x_discrete("Cross", labels = momdadcross.labels, drop=F) +
    theme_minimal()
  print(p)
}

traits %>% filter(generation=="F2") %>% select(data, yvar, ylab, type) %>% 
  pwalk(plot_F2, emm=F2.emm)
```

### Cytonuclear incompatibility

```{r f2plot_contrasts}
F2.contrasts <- tibble(momdadcross = c("H x HK","HK x H","H x KH","KH x H","K x KH","KH x K","K x HK","HK x K"),
                       pair = rep(1:4, each=2),
                       reciprocal.difference = rep(c("not expected", "expected"), times = 2, each = 2))

reorder_momdadcross <- function(df) {
  df %>% filter(momdadcross %in% F2.contrasts$momdadcross) %>% 
    left_join(F2.contrasts, by="momdadcross") %>% 
    mutate(momdadcross = fct_relevel(momdadcross, F2.contrasts$momdadcross)) 
}

plot_F2_contrasts <- function(emm, data, yvar, ylab, type="continuous") {
  data <- get(data) %>% reorder_momdadcross()
  emm <- emm %>% filter(trait==yvar) %>% reorder_momdadcross()
  sig <- F2.mod[[yvar]] %>% 
    multcomp::glht(linfct = multcomp::mcp(momdadcross = F2.contrasts %>% group_by(pair) %>% 
                                            summarize(contrast = paste0("`", paste(momdadcross[1:2], collapse="` - `"),"` == 0")) %>% 
                                            pull(contrast))) %>% summary(test=multcomp::adjusted("none")) %>% tidy() %>% mutate(pair=1:4)
  if(type=="continuous") {
    p <- ggplot(data, aes_string(y=yvar, x="momdadcross")) + 
      geom_violin(aes(fill=reciprocal.difference),draw_quantiles = (1:3)/4) + 
      geom_pointrange(data=emm, aes(y=estimate, ymax=conf.high, ymin=conf.low),color="black", show.legend = F) + 
      scale_fill_manual("Reciprocal difference     :::", values=brewer.pal(10,"Paired")[c(2,8)]) +
      scale_y_continuous(name = ylab, expand = expansion(mult=c(0, 0.05)), limits=c(0,NA))
  } else if(type=="binary") {
    p <- ggplot(data, aes_string(y=yvar, x="momdadcross")) + 
      geom_pointrange(data=emm %>% 
                        mutate(conf.low  = ifelse(conf.high-conf.low>0.9,NA,conf.low), 
                               conf.high = ifelse(conf.high-conf.low>0.9,NA,conf.high)), 
                      aes(y=estimate, ymax=conf.high, ymin=conf.low, color=reciprocal.difference)) + 
      scale_color_manual("Reciprocal difference  :", values=brewer.pal(10,"Paired")[c(2,8)])+
      scale_y_continuous(ylab, limits=c(0,1), labels=~scales::percent(.,accuracy=1))
  }
  ytop = ifelse(type=="binary", 0.95, 1.02*max(pull(data, yvar), na.rm=T))
  p <- p + scale_x_discrete("Cross", labels = momdadcross.labels) +
    theme_minimal() + theme(legend.position = "top") +
    geom_text(data=sig, aes(x = 2*pair-0.5, y=ytop*1.04, label=ifelse(p.value < 0.01, "< 0.01",round(p.value,2)))) +
    geom_segment(aes(x=2*pair-1, xend=2*pair, y=ytop, yend=ytop))
  print(p)
}

traits %>% filter(generation=="F2", data != "f1seedmass") %>% select(data, yvar, ylab, type) %>% 
  pwalk(plot_F2_contrasts, emm=F2.emm)
```

