---
title: "F1 Seed Mass GLMM"
author: "John Powers"
date: "Nov 14, 2017"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.main-container { max-width: 1600px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

#Purpose
Identify reproductive barriers between two sympatric moth-pollinated plant species, *Schiedea kaalae* and *S. hookeri* by fitting a generalized linear mixed model (GLMM).

Fixed effects:

* **crosstype**

Potential random effects:

* **mompid** - maternal plant, specified by its population and ID


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning= F, comment="  ", cache=F, fig.width=9)
library(lme4)
library(MASS)
library(car)
library(glmmADMB)
library(bbmle)
library(lsmeans)
library(vcd)
library(countreg)

source("glmm_funs.R")

#Visualization
library(coefplot2)
library(ggplot2); theme_set(theme_gray())
library(lattice)
library(plyr)
library(beanplot)
library(knitr)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

#Data Import

```{r data}
sds <- read.table("f1seedmass.csv", header=T, sep="\t")
sds$smass <- sds$smass/10

sdsm2 <- sds[grepl("x",sds$momfullcross),]
sdsm2.split <- cbind(sdsm2$crossid, as.data.frame(matrix(unlist(matrix(lapply(matrix(unlist(strsplit(as.character(sdsm2$momfullcross), " x ", fixed=T)), nrow=2), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=2)), ncol=4, byrow=T)))
colnames(sdsm2.split) <- c("crossid","mommompop","mommomid","momdadpop","momdadid")

sdsd2 <- sds[grepl("x",sds$dadfullcross),]
sdsd2.split <- cbind(sdsd2$crossid, as.data.frame(matrix(unlist(matrix(lapply(matrix(unlist(strsplit(as.character(sdsd2$dadfullcross), " x ", fixed=T)), nrow=2), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=2)), ncol=4, byrow=T)))
colnames(sdsd2.split) <- c("crossid","dadmompop","dadmomid","daddadpop","daddadid")

sdsm1 <- sds[!grepl("x",sds$momfullcross),]
sdsm1.split <- cbind(sdsm1$crossid, as.data.frame(matrix(unlist(matrix(lapply(as.character(sdsm1$momfullcross), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=2)), ncol=2, byrow=T)))
colnames(sdsm1.split) <- c("crossid","mompop","momid")

sdsd1 <- sds[!grepl("x",sds$dadfullcross),]
sdsd1.split <- cbind(sdsd1$crossid, as.data.frame(matrix(unlist(matrix(lapply(as.character(sdsd1$dadfullcross), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=2)), ncol=2, byrow=T)))
colnames(sdsd1.split) <- c("crossid","dadpop","dadid")


sds <- merge(sds, sdsm2.split, by="crossid", all.x=T)
sds <- merge(sds, sdsd2.split, by="crossid", all.x=T)
sds <- merge(sds, sdsm1.split, by="crossid", all.x=T)
sds <- merge(sds, sdsd1.split, by="crossid", all.x=T)

pop2sp <- function(x) { factor(ifelse(x %in% c("904","3587", "892"), "kaal", ifelse(is.na(x),NA,"hook")))}
sds$mommomspecies <- pop2sp(sds$mommompop)
sds$momdadspecies <- pop2sp(sds$momdadpop)
sds$dadmomspecies <- pop2sp(sds$dadmompop)
sds$daddadspecies <- pop2sp(sds$daddadpop)
sds$momcross <- factor(toupper(paste0(substr(sds$mommomspecies,0,1), substr(sds$momdadspecies,0,1))))
sds$dadcross <- factor(toupper(paste0(substr(sds$dadmomspecies,0,1), substr(sds$daddadspecies,0,1))))

sds$momcross <- factor(ifelse(sds$momcross=="NANA", toupper(substr(as.character(pop2sp(sds$mompop)),0,1)), as.character(sds$momcross)))
sds$dadcross <- factor(ifelse(sds$dadcross=="NANA", toupper(substr(as.character(pop2sp(sds$dadpop)),0,1)), as.character(sds$dadcross)))
is.na(sds$dadcross) <- is.na(sds$momcross) <- sds$crosstype=="control"

table(sds$crossF)
with(sds, beanplot(smass~crossF, las=2))

sds$dadcross <- factor(sds$dadcross, c("H",  "HK", "KH", "K"))
sds$momcross <- factor(sds$momcross, c("H",  "HK", "KH", "K"))

table(sds$momcross, sds$dadcross)
table(sds$momcross, sds$dadcross, sds$crossF) #errors in crosstype!


#table(sds$momcross, sds$dadcross)

#sds <- sds[!(sds$momcross=="KH" & sds$dadcross=="HK"),]
```

#Models

```{r models, echo=T}
f1.md        <- glm(smass~momcross*dadcross, data=sds, family="gaussian")
```

#Model summaries
```{r summaries}
summary(f1.md)
AIC(f1.md)
Anova(f1.md)
```

#lsmeans and Tukey tests

```{r lsmeanstr}
#Count
rg.tr <- ref.grid(f1.md)
#summary(rg)
c.lsm <- lsmeans(rg.tr, ~ momcross*dadcross)
plot(c.lsm)
cld.mod <- cld(c.lsm, Letters=letters) #tukey letterings
cld.mod$response <- cld.mod$lsmean
cld.mod$uSE <- cld.mod$lsmean+cld.mod$SE
cld.mod$lSE <- cld.mod$lsmean-cld.mod$SE
cld.mod

ggplot(as.data.frame(cld.mod), aes(y=response, x=momcross, fill=dadcross)) +
  geom_col(position=position_dodge2()) +
  geom_linerange(aes(ymin=lSE, ymax=uSE), position=position_dodge2(0.9)) +
  labs(x="Maternal plant", y="Seed mass (g)") +
  geom_text(aes(label=trimws(.group), x=momcross, hjust=length(trimws(.group))/30), position=position_dodge2(0.9), vjust=-2) + 
  geom_text(aes(label=dadcross, y=0.05), position=position_dodge2(0.9)) +
  scale_y_continuous(expand = expand_scale(mult=c(0,.1)), breaks = scales::pretty_breaks(n = 10))+
  scale_fill_manual(values=c("gray80","gray60","gray20","gray40"))+
  theme_classic() + theme(legend.text=element_text(size=rel(1)), legend.position="bottom", axis.text = element_text(colour="black", size=rel(1)), text=element_text(size=14), axis.ticks.x = element_blank()) + guides(fill=F)
```
