---
title: "Backcross Pollen vs. Backcross Seeds"
author: "John Powers"
date: "Dec 9, 2019"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.table { width: 40%; }
</style>

Modified from hybridViability.Rmd (Oct 2017)

# Purpose
Test for trade-offs between male and female function in F2 hybrids of two sympatric moth-pollinated plant species, *Schiedea kaalae* and *S. hookeri*.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning= F, comment="  ", cache=T)
library(lme4)
library(MASS)
library(car)
library(glmmADMB)
library(bbmle)
library(emmeans)
library(vcd)
library(countreg)

source("glmm_funs.R")

# Visualization
library(coefplot2)
library(ggplot2); theme_set(theme_minimal())
library(lattice)
library(plyr)
library(beanplot)
library(knitr)
library(dplyr)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

# Data Import

```{r data}
setwd("~/MyDocs/MEGA/UCI/Schiedea/Analysis/HybridSeeds")
pol <- read.table("f2pollenEH.csv", header=T, sep="\t", colClasses=c(date="Date"))
pol <- pol[!is.na(pol$V1) & pol$dadfullcross!="9999",]

pol$v <- rowSums(pol[,grepl("V", colnames(pol))])/10 # average viable counts
pol$i <- rowSums(pol[,grepl("I", colnames(pol))])/10 # average inviable counts
pol$vpf <- (pol$v * 444.4/20) * 10      # number viable per flower (5 flrs)
pol$ipf <- (pol$i * 444.4/20) * 10      # number inviable per flower (5 flrs)
pol$vp <- pol$vpf / (pol$vpf + pol$ipf) # proportion viable
pol$tpf <- pol$vpf + pol$ipf
pol$t <- 10*(pol$v + pol$i)

polm2 <- pol[grepl("x",pol$momfullcross),]
polm2.split <- cbind(polm2$crossid, as.data.frame(matrix(unlist(matrix(lapply(matrix(unlist(strsplit(as.character(polm2$momfullcross), " x ", fixed=T)), nrow=2), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=2)), ncol=4, byrow=T)))
colnames(polm2.split) <- c("crossid","mommompop","mommomid","momdadpop","momdadid")

pold2 <- pol[grepl("x",pol$dadfullcross),]
pold2.split <- cbind(pold2$crossid, as.data.frame(matrix(unlist(matrix(lapply(matrix(unlist(strsplit(as.character(pold2$dadfullcross), " x ", fixed=T)), nrow=2), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=2)), ncol=4, byrow=T)))
colnames(pold2.split) <- c("crossid","dadmompop","dadmomid","daddadpop","daddadid")

polm1 <- pol[!grepl("x",pol$momfullcross),]
polm1.split <- cbind(polm1$crossid, as.data.frame(matrix(unlist(matrix(lapply(as.character(polm1$momfullcross), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=1)), ncol=2, byrow=T)))
colnames(polm1.split) <- c("crossid","mompop","momid")

pold1 <- pol[!grepl("x",pol$dadfullcross),]
pold1.split <- cbind(pold1$crossid, as.data.frame(matrix(unlist(matrix(lapply(as.character(pold1$dadfullcross), function(x) { spl <- unlist(strsplit(unlist(strsplit(x, " ", fixed=T)), "-", fixed=T)); c(spl[1],gsub("-NA","",paste(spl[2],spl[3],spl[4], sep="-"))) }), ncol=1)), ncol=2, byrow=T)))
colnames(pold1.split) <- c("crossid","dadpop","dadid")


pol <- merge(pol, polm2.split, by="crossid", all.x=T)
pol <- merge(pol, pold2.split, by="crossid", all.x=T)
pol <- merge(pol, polm1.split, by="crossid", all.x=T)
pol <- merge(pol, pold1.split, by="crossid", all.x=T)

popcols <- c("mompop","dadpop","mommompop","momdadpop","dadmompop","daddadpop")
pol[popcols] <- lapply(pol[popcols], function(x) { sapply(x, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"), warn_missing=F) })

kaalpops <- c("904WPG","3587WP", "892WKG")
pop2sp <- function(x) { factor(ifelse(x %in% kaalpops, "kaal", ifelse(is.na(x),NA,"hook")))}
pol$mommomspecies <- pop2sp(pol$mommompop)
pol$momdadspecies <- pop2sp(pol$momdadpop)
pol$dadmomspecies <- pop2sp(pol$dadmompop)
pol$daddadspecies <- pop2sp(pol$daddadpop)
pol$momcross <- factor(toupper(paste0(substr(pol$mommomspecies,0,1), substr(pol$momdadspecies,0,1))))
pol$dadcross <- factor(toupper(paste0(substr(pol$dadmomspecies,0,1), substr(pol$daddadspecies,0,1))))

pol$momcross <- factor(ifelse(pol$momcross=="NANA", toupper(substr(as.character(pop2sp(pol$mompop)),0,1)), as.character(pol$momcross)))
pol$dadcross <- factor(ifelse(pol$dadcross=="NANA", toupper(substr(as.character(pop2sp(pol$dadpop)),0,1)), as.character(pol$dadcross)))
# is.na(pol$dadcross) <- is.na(pol$momcross) <- pol$crosstype=="control"

pol$dadcross <- factor(pol$dadcross, c("H",  "HK", "KH", "K"))
pol$momcross  <- factor(pol$momcross, c("H",  "HK", "KH", "K"))
pol$crosstype <- with(pol, factor(paste(momcross, dadcross, sep="x")))
pol$species <- factor(ifelse(pol$mompop %in% kaalpops, "kaal", "hook"))
pol$dadsp <-   factor(ifelse(pol$dadpop %in% kaalpops, "kaal", "hook"))
pol <- within(pol, smompop <- as.factor(paste(species,mompop,sep="")))
pol <- within(pol, mompid <- as.factor(paste(mompop,momid,sep=".")))
pol <- within(pol, dadpid <- as.factor(paste(dadpop,dadid,sep=".")))

pol$mompop <- as.character(pol$mompop)
pol$mompop[is.na(pol$mompop)] <- as.character(pol$momcross[is.na(pol$mompop)])
pol$mompop <- as.factor(pol$mompop)
pol$dadpop <- as.character(pol$dadpop)
pol$dadpop[is.na(pol$dadpop)] <- as.character(pol$dadcross[is.na(pol$dadpop)])
pol$dadpop <- as.factor(pol$dadpop)

```

```{r f2seeddata}
setwd("~/MyDocs/MEGA/UCI/Schiedea/Analysis/HybridSeeds")
sds <- read.table("f2seedsEH.csv", header=T, sep="\t", colClasses=c(date_emasculation="Date",	date_pollination="Date"))

msds <- sds %>% 
  group_by(crossid,pollen_donor) %>% 
  dplyr::summarize(viable_seeds = mean(viable_seeds), inviable_seeds = mean(inviable_seeds)) %>%
  mutate(pviable_seeds = viable_seeds/(viable_seeds+inviable_seeds))

sdspol <- merge(msds,pol,by="crossid")

str(sdspol)
```

# Inventory
Take out F2s before analysis
```{r inventory}
table(sdspol$crosstype)
sdspol <- sdspol %>% filter(!(crosstype %in% c("HKxHK","KHxKH")))
```

# Analyses

## Viable seeds vs. viable pollen
```{r sdspol_plot, fig.height=7}
(sdspol.plot <- ggplot(sdspol, aes(x=vpf, y=viable_seeds, color=crosstype, linetype=pollen_donor, shape=pollen_donor)) + 
   geom_point(size=3) +
   #geom_smooth(method = "lm", se=F, aes(group=crosstype)) + 
   # geom_quantile(aes(group=1))+
   xlab("Viable pollen grains per flower") + 
   ylab("Mean number of viable seeds") +
   scale_color_brewer("Backcross Type", type="qual", palette="Paired") +
   #scale_shape_manual("Maternal Plant", values=c(15, 18, 16, 17))+  
   scale_shape_manual("Pollen donor", values=c(15, 19))+
   theme_minimal()+
   theme(legend.key.size = unit(1.7,"lines")))
#ggsave("viableseeds_viablepollen.png", sdspol.plot, type="cairo", width=6, height=6)
```

Variables: viable pollen grains per flower, mean number of viable seeds.

1. Test for an overall relationship at the crosstype level, taking the mean values for each crosstype (low N)

```{r sdspol_mean_model, echo=TRUE}
sdspol.mean <- sdspol %>% group_by(crosstype) %>% summarize_at(c("viable_seeds", "vpf"), mean)
sdspol.mean.mod <- lm(viable_seeds ~ vpf, data=sdspol.mean)
summary(aov(sdspol.mean.mod))
```

2. Test for an overall relationship at the plant level, ignoring crosstype

```{r sdspol_overall_model, echo=TRUE}
sdspol.overall.mod <- lm(viable_seeds ~ vpf, data=sdspol)
summary(aov(sdspol.overall.mod))
```

3. Test for a relationship within each cross type 

```{r sdspol_model, echo=TRUE}
sdspol.mod <- lm(viable_seeds ~ vpf * crosstype, data=sdspol)
summary(aov(sdspol.mod))
summary(sdspol.mod)
```


```{r sdspol_percent_plot, fig.height=7, include=F}
(sdspol.percent.plot <- ggplot(sdspol, aes(x=vp, y=pviable_seeds, color=crosstype, linetype=momcross, shape=pollen_donor)) + 
   geom_point(size=3) +
   #geom_smooth(method = "loess", se=F, span=2, aes(group=momcross)) + 
   xlab("Percent viable pollen") + 
   ylab("Percent viable seeds") +
   scale_x_continuous(labels = scales::percent_format(accuracy=1)) +
   scale_y_continuous(labels = scales::percent_format(accuracy=1)) +
   scale_color_brewer("Backcross Type", type="qual", palette="Paired") +
   #scale_shape_manual("Maternal Plant", values=c(15, 18, 16, 17))+
   scale_shape_manual("Pollen donor", values=c(15, 19))+
   theme_minimal()+
   theme(legend.key.size = unit(1.7,"lines")))
#ggsave("viableseeds_viablepollen.png", sdspol.plot, type="cairo", width=6, height=6)
```

```{r sdspol_percent_model, include =F}
sdspol.percent.mod <- lm(pviable_seeds ~ vp * crosstype + pollen_donor, data=sdspol)
summary(aov(sdspol.percent.mod))
```
