---
title: "Hybrids Inflo Biomass GLMM"
author: "John Powers"
date: "October 15, 2017"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.table { width: 40%; }
</style>

Modified from hybridInflo Biomass.Rmd (April 2017)

#Guides
* Bolker et al. 2009. [Generalized linear mixed models: a practical guide for ecology and evolution](http://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(09)00019-6). Trends in Ecology & Evolution.
* Bolker. [GLMM FAQ](http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html)
* Bolker et al. 2011. GLMMs in action: gene-by-environment interaction in total fruit production of wild populations of *Arabidopsis thaliana*. [Part 1](http://www.cell.com/cms/attachment/601623/4742452/mmc1.pdf), [Part 2](http://www.cell.com/cms/attachment/601623/4742453/mmc2.pdf)
* Bolker. In: Ecological Statistics: Contemporary Theory and Application
     + [GLMM:worked examples](http://ms.mcmaster.ca/~bolker/R/misc/foxchapter/bolker_chap.html) 
     + [Reporting GLMM results](http://ms.mcmaster.ca/~bolker/misc/GLMM_results_report.pdf)
* Bolker et al. 2013. [Strategies for fitting nonlinear ecological models in R, AD Model Builder, and BUGS](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12044/full)
     + Bolker et al. 2012. [Owls example: a zero-inflated, generalized linear
mixed model for count data](https://groups.nceas.ucsb.edu/non-linear-modeling/projects/owls/WRITEUP/owls.pdf)
     + other [examples](https://groups.nceas.ucsb.edu/non-linear-modeling/) from Non-Linear Model Working Group
* [A Practical Guide to Mixed Models in R](http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html)
* Schielzeth and Nakagawa 2012. [Nested by design: model fitting and interpretation in a mixed model era](http://onlinelibrary.wiley.com/doi/10.1111/j.2041-210x.2012.00251.x/full). Methods in Ecology and Evolution.
* Stroup. [Rethinking the Analysis of Non-Normal Data in Plant and Soil Science](https://dl.sciencesocieties.org/publications/aj/articles/107/2/811)
* Elston et al. 2001. [Analysis of aggregation, a worked example: Numbers of ticks on red grouse chicks](https://www.researchgate.net/publication/11945272_Analysis_of_aggregation_a_worked_example_Numbers_of_ticks_on_red_grouse_chicks)
* Annika Nelson. [Ant GLMM Example](http://rpubs.com/annikanelson/GLMM_Example)
* R package *pscl* vignette: [Regression Models for Count Data in R](https://cran.r-project.org/web/packages/pscl/vignettes/countreg.pdf)

#Purpose
Identify reproductive barriers between two sympatric moth-pollinated plant species, *Schiedea kaalae* and *S. hookeri* by fitting a generalized linear mixed model (GLMM).

In the experimental design, the following **crosstypes** were made:

* within species, **between** population (may show outbreeding depression or heterosis)
* within species, **within** populations (may show inbreeding depression)
* **hybrids** between species (indicates species barrier from pollination to seed production)

In this analysis the response variable is the biomass of the offspring produced by each cross. Other barriers (hybrid survival, flowering) could be analyzed in a similar framework, with appropriate changes to the underlying distribution. 

Fixed effects:

* **crosstype** - hybrids, within population, between populations
* **species** - species of the maternal plant that produced the Inflo Biomass

Potential random effects:

* **mompop** - maternal plant population
* **mompid** - maternal plant, specified by its population and ID
* **dadpop** - paternal plant population

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning= F, comment="  ", cache=T)
library(lme4)
library(MASS)
library(car)
library(glmmADMB)
library(bbmle)
library(lsmeans)
library(vcd)
library(countreg)

source("glmm_funs.R")

#Visualization
library(coefplot2)
library(ggplot2); theme_set(theme_gray())
library(lattice)
library(plyr)
library(beanplot)
library(knitr)
library(RColorBrewer)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

#Data Import

```{r data}
ib <- read.table("inflobiomass.csv", header=T, sep="\t", 
                 colClasses=c(collect.date="Date", weigh.date="Date"))
crosses <- read.table("hybrids.csv", header=T, sep="\t", colClasses=c(mompop="factor", dadpop="factor"))

crosscol <- c("green","blue","orange","red")

#treat populations as factors
ib$mompop <- crosses$mompop[match(ib$crossid, crosses$crossid)]
ib$momid <- crosses$momid[match(ib$crossid, crosses$crossid)]
ib$species <- crosses$momsp[match(ib$crossid, crosses$crossid)]
ib$dadpop <- crosses$dadpop[match(ib$crossid, crosses$crossid)]
ib$dadid <- crosses$dadid[match(ib$crossid, crosses$crossid)]
ib$dadsp <- crosses$dadsp[match(ib$crossid, crosses$crossid)]
ib$crosstype <- crosses$crosstype[match(ib$crossid, crosses$crossid)]
ib$cross <- crosses$cross[match(ib$crossid, crosses$crossid)]

#rename crosstype codes
ib$crosstype <- factor(ib$crosstype, levels=c("between", "within", "hybrid"))
#made "between" the first reference level to facilitate comparison between outcrossing populations and hybridizing species 

ib$mompop <- sapply(ib$mompop, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"))
ib$dadpop <- sapply(ib$dadpop, mapvalues, from = c("794","866","899","879","892","904","3587"), to = c("WK","WK","WK","879WKG","892WKG","904WPG","3587WP"))

#define interactions
ib <- within(ib, sxc <- interaction(species,crosstype))
ib <- within(ib, sxcxm <- interaction(species,crosstype,mompop,momid))
ib <- within(ib, mompid <- as.factor(paste(mompop,momid,sep=".")))
ib <- within(ib, dadpid <- as.factor(paste(dadpop,dadid,sep=".")))
ib <- within(ib, smompop <- as.factor(paste(species,mompop,sep="")))

ib$collect.date <- round(difftime(ib$collect.date, "2016-03-10"))
#check final structure
str(ib)


```

#Data Inspection

##Regression
```{r regression, warning=FALSE}
library(ggplot2)
qplot(log10(mass), log10(flrs), col=cross, data=ib, weight=mass) + geom_smooth(method="lm", se=T) + scale_color_manual(values=crosscol) + ggtitle("cross")

qplot(log10(mass), log10(flrs), col=cross, group=paste(mompop,dadpop), data=ib) +  geom_smooth(method="lm", se=F) + scale_color_manual(values=crosscol) + ggtitle("mompop*dadpop")

#qplot(log10(mass), log10(flrs), col=cross, data=ib, weight=mass) + geom_quantile(quantiles = 0.9) + scale_color_manual(values=crosscol)
  
#plot(log10(flrs)~log10(mass), data=ib, type="n")
#text(log10(ib$mass), log10(ib$flrs), 1:length(ib$plantid), col=crosscol[ib$cross])

#library(quantreg,log10(flrs)~log10(mass)*cross, data=ib[-exclude,], weights=mass)
#qr <- rq(log10(flrs)~log10(mass)*cross, data=ib[-exclude,], tau = 0.9)
#summary(qr)

mf <- lm(log10(flrs)~log10(mass)*cross, data=ib, weights=mass)
plot(mf)
mass.flrs <- summary(mf)
mass.flrs

slopes <- mass.flrs$coefficients[c(2,6,7,8),"Estimate"]
slopes[c(2,3,4)] <- slopes[c(2,3,4)]+slopes[1]
names(slopes) <- levels(ib$cross)
slopes

intercepts <- mass.flrs$coefficients[c(1,3,4,5),"Estimate"]
intercepts[c(2,3,4)] <- intercepts[c(2,3,4)]+intercepts[1]
names(intercepts) <- levels(ib$cross)
intercepts

#with(ib, plot(mass,flrs, col=crosscol[cross], xlim=c(0,0.5), ylim=c(0,240)))
#for(i in 1:4) { curve(10^intercepts[i]*x^slopes[i], from=0, to=6, col=crosscol[i], add=T) }
#with(ib, plot(log10(mass),log10(flrs), col=crosscol[cross]))
#for(i in 1:4) { curve(intercepts[i]+x*slopes[i], from=-2, to=1, col=crosscol[i], add=T) }

```

```{r addr}
#Add together envelopes for regression and the rest
ibold <- ib

ib.inflos <- aggregate(inflo.e ~ crossid+plantid+cross+mompop+momid+species+dadpop+dadid+dadsp+crosstype+sxc+sxcxm+mompid+dadpid+smompop, ibold, sum, na.action=na.pass)
ib.inflos.mean <- aggregate(inflo.e ~ crosstype+species, ib.inflos, mean, na.action=na.omit)

ib <- aggregate(mass ~ crossid+plantid+cross+mompop+momid+species+dadpop+dadid+dadsp+crosstype+sxc+sxcxm+mompid+dadpid+smompop, ib, sum, na.action=na.pass)
ib$collect.date <- aggregate(collect.date ~ crossid+plantid+cross+mompop+momid+species+dadpop+dadid+dadsp+crosstype+sxc+sxcxm+mompid+dadpid+smompop, ibold, max, na.action=na.pass)$collect.date
```

```{r datemass, warning=FALSE}
library(ggplot2)
ggplot(ib, aes(x=log10(mass), fill=cross, color=cross)) + geom_density(alpha=0.1) + scale_fill_manual(values=crosscol) +scale_color_manual(values=crosscol)
```

##Replication
The sample sizes are unbalanced at all levels, including maternal population: 
```{r reps}
reptab <- with(ib, table(smompop,crosstype))
mosaic(reptab, pop=F)
labeling_cells(text = reptab, margin = 0)(reptab)
```

Replication is low for some **within**-population crosses. The replication is even lower for each maternal plant, so we need to be wary of estimates when subsetting at this level:

```{r reps_mpid}
with(ib, kable(table(mompid,crosstype)))
```

##Overall data distribution

To identify the best-fitting distribution, we make quantile-quantile plots of the raw data against various distributions. The more points within the confidence interval envelopes, the better the fit. Later, we present quantile-quantile plots of the model residuals to assess model fit.

```{r dists}
#QQ plots against various distributions
set.seed(1)
par(mfrow=c(1,3))
normal <- fitdistr(log10(ib$mass+1), "normal")
qqp(log10(ib$mass+1), "norm", main="Normal")
lognormal <- fitdistr(ib$mass+1, "lognormal")
qqp(ib$mass+1, "lnorm", main="Log Normal")
#pois <- fitdistr(ib$mass+1, "Poisson")
#qqp(ib$mass, "pois", pois$estimate, main="Poisson")
#nbinom <- fitdistr(ib$mass+1, "Negative Binomial") 
#qqp(ib$mass+1, "nbinom", size = nbinom$estimate[[1]], mu=nbinom$estimate[[2]], main="Negative Binomial")
gamma <- fitdistr(ib$mass+1, "gamma")
qqp(ib$mass+1, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]], main="Gamma")
```

##Distributions by fixed factors
```{r hst_fixed}
ggplot(ib, aes(x = mass, fill=species)) +
  geom_histogram(data=subset(ib,species == "hook"), aes(y=-..density..),binwidth=1)+
  geom_histogram(data=subset(ib,species == "kaal"), aes(y= ..density..),binwidth=1)+
  coord_flip() + facet_grid(~crosstype) + labs(y="Histogram", x="Inflo Biomass")
```

##Distributions by random factors
```{r hst_rand_mpid}
ggplot(aes(y=mass, x=mompid, color=crosstype), data=ib) + geom_count(alpha=0.8) + coord_flip() + labs(x="Maternal plant", y="Mass")
```

##Homogeneity of variances across subsets
Our mixed model uses one parameter to capture random effect variance, which is assumed to be homogeneous. Plotting on a log scale should uncouple variances from means to assess this visually. Subsets are **species** \* **crosstype** \* **maternal plant**.

Subset variances are not homogeneous:
```{r varhom}
ggplot(aes(y=log10(mass+1), x=sxcxm, color=crosstype), data=ib) + geom_boxplot() + coord_flip() + labs(y="ln(Inflo Biomass + 1)",x="Subsets")
```


##Subset mean-variance relationship
Various distributions make different assumptions about the mean-variance (µ-Var) ratio. 

```{r meanvar}
grpVars <- with(ib, tapply(mass, list(sxcxm), var))
grpMeans <- with(ib, tapply(mass, list(sxcxm), mean))
grpCounts <- with(ib, tapply(mass, list(sxcxm), length))
#set weight=grpCounts to weight loess by sample sizes
ggplot(na.omit(data.frame(grpMeans,grpVars,grpCounts)),
       aes(x=grpMeans,y=grpVars, weight=1))+geom_point(aes(size=grpCounts))+
  guides(colour=guide_legend(title="Fit"),size=guide_legend(title="Sample size")) + labs(x="Subset Mean", y="Subset Variance") + labs(subtitle="Subset: species*crosstype*mompid")
```

##Fixed effects
Effects and interactions in these plots are simply given by the mean, which may be unduly influenced by high values.
```{r maineff}
intplot <- ggplot(ib,aes(x=crosstype,y=mass))+
  geom_count(aes(size = ..prop.., group=sxc),alpha=0.5)+
  stat_summary(aes(x=as.numeric(crosstype)),fun.y=mean,geom="line")+ facet_grid(~species)
intplot + aes(group=species, color=species)
```

##Random effects

### Maternal population
```{r randeff_mp}
intplot + aes(group=mompop, color=mompop)
```

### Maternal plant
```{r randeff_mpid}
intplot + aes(group=mompid, color=mompop)
```

### Paternal population
```{r randeff_dp}
intplot + aes(group=dadpop, color=dadpop)
```


##Run models on subsets
Run many generalized linear models on subsets of the data defined by **crosstype | mompid** to see if effects estimates are consistent within maternal plants.

Most maternal plant subsets agree, but some are problematic outliers. These plants can be picked out visually from the random effects interaction plot above, the estimated parameters of each subset model, and the QQ plot of the estimated parameters:

```{r lmlist, results=FALSE, warning=FALSE}
 #had to get rid of species or mompid since mompid is nested inside species. dadpop also works
glm.lis <- lmList(log10(mass)~crosstype|mompid,data=ib, family="gaussian")
plot.lmList(glm.lis,scale=list(x=list(relation="free")))
```

```{r qqmath, results=FALSE, warning=FALSE}
qqmath.lmList(glm.lis)#
```


#Models
We constructed the following models with the package *glmmADMB*. They all have the same fixed effects, **species x crosstype**, and response variable, **log10(mass)**

* X  = standard GL(M)M

| Distribution, Random Effects: | None | Maternal plant | Maternal population |
|-------------------------------|------|----------------|---------------------|
| normal (norm) | X | X | X |

```{r models, echo=T, cache=T}
#Normal (Gaussian) distribution, identity link
ib <- ib[!is.na(ib$collect.date),]
sc.norm.l          <- lm(log10(mass)~species*crosstype+collect.date, data=ib)
sc.mix.mompid.l           <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompid), data=ib)
sc.nd.mix.momdadpid.l           <- lmer(log10(mass)~species*crosstype + (1|mompid)+ (1|dadpid), data=ib)
sc.mix.mompop.l    <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompop), data=ib)
sc.mix.momdadpid.l           <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompid) + (1|dadpid), data=ib)
sc.mix.momdadpop.l    <- lmer(log10(mass)~species*crosstype+collect.date + (1|mompop) + (1|dadpop), data=ib)
```

#Model comparison
##AIC
We will use the Aikake Information Criterion to pick the model the best fits the data, penalized by the number of parameters. Differences of 2 units are significant.


```{r aic}
sc.names <- c("sc.norm.l","sc.mix.mompid.l","sc.mix.mompop.l","sc.mix.momdadpid.l","sc.mix.momdadpop.l")
sc.list <- sapply(sc.names, get, USE.NAMES=T)
sc.AIC <- ICtab(sc.list,mnames=sc.names,type="AIC", base=T, delta=F) # for AICc, nobs=nobs(sc.list[[1]])
class(sc.AIC)<-"data.frame"
all.names <- c(sc.names)
all.list <- sapply(all.names, get, USE.NAMES=T)
all.AIC <- dfun(rbind(sc.AIC))
all.AIC <- all.AIC[order(all.AIC$dAIC),]
kable(all.AIC, format.arg=list(digits=3))
```

The best-fiting model is a mixed model with the following components:

* 
     + response: **log10(mass)**
     + fixed effects: **species**, **crosstype**, **species x crosstype**
     + random effect: **mompid**


##Overdispersion

Looking at the normal, fixed effects model, we see that the residuals are not normal:
```{r shapiro}
shapiro.test(sc.norm.l$residuals)#raw residuals!
```

##Coefficients
The coefficients estimated for each model agree qualitatively.


```{r coef, fig.asp=1}
sc.log.names <- sc.names
sc.log <- sapply(sc.log.names, get, USE.NAMES=T)

coefplot2(sc.log, legend.x="topright",legend=T,legend.args=list(cex=0.8, xpd=T, inset=c(-0.1,0)), col.pts=sample(gg_color_hue(length(sc.log.names))), spacing=0.05, lwd.2=2, lwd.1=4, intercept=F)
```

#Inference
We chose the model with nearly the best (lowest) AIC, to carry out inference tests and parameter estimation.

##Description


```{r pick}
mod <- sc.mix.momdadpid.l
print(mod)
```

##Test significance of random effects
Using a likelihood ratio test, with a null hypothesis of zero variance, the random effect (maternal plant) is significant for both model parts:


```{r testrand, eval=FALSE}
anova(sc.norm.l, sc.mix.mompid.l) #double this p-value. or simulate null by permuting data.
```

##Test significance of interaction
By dropping it from the model and performing a likelihood-ratio test, we see that the **species x crosstype** interaction is significant for the count model but not the binary model:


```{r testint}
sxc.chisq <- drop1(mod, test="Chisq") #load from file
dfun(sxc.chisq)
```

##Model summary
The model estimated the following parameters, with individual parameter significance determined by the Wald z-test, and fixed effect significance determined by analysis of deviance Wald test.


```{r summ}
summary(mod)
Anova(mod, type=3)
```

##Predicted random effects
These are box and QQ (to check normality) plots of the estimated random effect of each maternal plant. 


```{r fitrand, eval=FALSE}
predre <- setNames(data.frame(RE=ranef(mod)$mompid,SD=ranef(mod, sd=T)$`1`),c("RE","SD"))
ggplot(predre, aes(x = rownames(predre),y=RE)) +
  geom_point(size = 2) + coord_flip()+
  geom_errorbar(aes(ymin = RE-SD, ymax = RE+SD)) + labs(x="Maternal plants", y="Predicted random effects")

#Count
reStack <- ldply(ranef(mod))
print( qqmath( ~`(Intercept)`|.id, data=reStack, scales=list(relation="free"),
                 prepanel = prepanel.qqmathline,
                 panel = function(x, ...) {
                   panel.qqmathline(x, ...)
                   panel.qqmath(x, ...)
                 },
                 layout=c(1,1)))
```

##Least square means
The least square means procedure can generate predictor estimates of each type, and give their significance groupings with a post-hoc Tukey test. *S. hookeri*-produced hybrids produce less Inflo Biomass than either crosses between or within *S. hookeri* populations. The other differences are not significant, but remember that the fixed effect of hybrid (vs. between) was significant (model summary).


```{r lsmeans, fig.asp=1, fig.width=5}
#Count
rg <- ref.grid(sc.mix.momdadpid.l)
#summary(rg)
sxc.lsm <- lsmeans(rg, ~ crosstype*species)
plot(sxc.lsm)
options(digits=4)
cld.mod <- cld(sxc.lsm, Letters=letters) #tukey letterings
cld.mod$response <- 10 ^ cld.mod$lsmean
cld.mod$uSE <- 10 ^ (cld.mod$lsmean+cld.mod$SE)
cld.mod$lSE <- 10 ^ (cld.mod$lsmean-cld.mod$SE)
cld.mod[rev(order(cld.mod$species, cld.mod$crosstype)),]

H.wb <-  with(cld.mod[cld.mod$species=="hook",], response[crosstype=="within"]/response[crosstype=="between"] - 1)
K.wb <- with(cld.mod[cld.mod$species=="kaal",], response[crosstype=="within"]/response[crosstype=="between"] - 1)
K.H <- with(cld.mod[cld.mod$crosstype=="between",], response[species=="kaal"]/response[species=="hook"] - 1)
maxsp <- ifelse(K.H>0, "kaal","hook")
minsp <- ifelse(K.H<0, "kaal","hook")
maxresp <- with(cld.mod, response[species==maxsp & crosstype=="between"])
minresp <- with(cld.mod, response[species==minsp & crosstype=="between"])
HK.resp <-  with(cld.mod, response[species=="hook" & crosstype=="hybrid"])
KH.resp <-  with(cld.mod, response[species=="kaal" & crosstype=="hybrid"])
HK.int <-   with(cld.mod, ifelse(HK.resp > minresp & HK.resp < maxresp, (HK.resp-minresp)/(maxresp-minresp), 
                                 ifelse(HK.resp < minresp, HK.resp/minresp-1, HK.resp/maxresp-1)))
KH.int <-   with(cld.mod, ifelse(KH.resp > minresp & KH.resp < maxresp, (KH.resp-minresp)/(maxresp-minresp), 
                                 ifelse(KH.resp < minresp, KH.resp/minresp-1, KH.resp/maxresp-1)))

intermed <- (minresp + maxresp) / 2
with(ib, wilcox.test(ib[species=="kaal" & crosstype=="hybrid","mass"], mu=intermed))
with(ib, wilcox.test(ib[species=="hook" & crosstype=="hybrid","mass"], mu=intermed))

round(c(H.wb,K.wb,K.H,HK.int,KH.int),2)

ggplot(as.data.frame(cld.mod), aes(y=response, x=relevel(crosstype, "within"), fill=species)) +
  geom_col(position=position_dodge2()) +
  geom_linerange(aes(ymin=lSE, ymax=uSE), position=position_dodge(0.9)) +
  labs(x="", y="Inflorescence biomass (g)",fill="Maternal species") +
  scale_fill_manual(labels = c("S. hookeri  ", "S. kaalae  "), values=brewer.pal(name="Set1", n=3)[c(3,2)]) +
  scale_x_discrete(labels = c("Intrapopulation", "Interpopulation", "Hybrid")) +
  geom_text(aes(label=.group), position=position_dodge(0.9), hjust=0, vjust=-1) +
  scale_y_continuous(expand = expand_scale(add=c(0,0)), breaks = scales::pretty_breaks(n = 5)) +
  theme_classic() + theme(legend.text=element_text(face="italic", size=rel(1)), legend.position="bottom", axis.text = element_text(colour="black", size=rel(1)), text=element_text(size=14), axis.ticks.x = element_blank()) + geom_segment(aes(x=2.5, y=intermed, xend=3.5, yend=intermed))
```


